---
title: "MM Avatar cleaning"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cerulean #darkly
    highlight: pygments
    df_print: paged
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   <!-- margin-bottom: 100px; -->
}

table {
    margin-top: 25px;
    <!-- margin-bottom: 100px !important; -->
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")

```

<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">Workflow through the MM Avatar data cleaning</span>

</div>
<br>

```{r library, include=FALSE}
library(tidyverse)
library(data.table)
library(lubridate)
```

***
# I.Clinical data

```{r load}
#######################################################################################  I  ### Load data----
path <- "/Users/colinccm/Documents/GitHub/Gillis/no vpn data/CHIP in Avatar"
# 1.1.Load Demographics data -------------------------------------------------------------------------------------
Demo_RedCap_V4ish <-
  readxl::read_xlsx(paste0(path, "/Raghu MM/extracted Avatar V124 data and dict/Avatar_Demographics_All_MM_modif_06292020.xlsx")) %>%
  select(c("avatar_id","TCC_ID","Date_of_Birth", "Gender", "Ethnicity", "Race"))
Demo_HRI <- 
  readxl::read_xlsx(paste0(path, "/Raghu MM/extracted Avatar V124 data and dict/Demographics_HRI_Export.xlsx")) %>%
  select(c("MRN",Date_of_Birth = "Date of Birth", Gender = "Gender Cerner", Ethnicity = "Ethnicity Cerner", Race = "Race Cerner"))
Demo_linkage <- 
  readxl::read_xlsx(paste0(path, "/Raghu MM/extracted Avatar V124 data and dict/Demographics_HRI_Export.xlsx"),
                    sheet = "Sheet1")
# 1.2.Load Germline with disease status --------------------------------------------------------------------------
Germline <- readxl::read_xlsx(paste0(path, 
                                     "/Raghu MM/Germline data/Moffitt_Germl_v0.4.5_Disease_Classification_12082020.xlsx")) %>% 
  select(c("avatar_id", "SLID_germline", "collectiondt_germline", 
           "moffittSampleId_germline", Disease_Status_germline = "Disease_Status"))
# uid <- paste(unique(Germline$avatar_id), collapse = '|')
# Germ <- 
#   readxl::read_xlsx(paste0(path, 
#                            "/Raghu MM/Germline data/Moffitt_Germl_v0.4.3_Disease_Classification_OUT01312020.xlsx")) %>% 
#   select(c("avatar_id", "collectiondt", "WES_HUDSON_ALPHA", "Disease_Status"))
# Germ <- Germ[(!grepl(uid, Germ$avatar_id)),]
# Germ2 <-
#   readxl::read_xlsx(paste0(path,
#                            "/Raghu MM/Germline data/Moffitt_Germl_Disease_Classification_2patient_from_2nd_sequencingfile.xlsx")) %>%
#   select(-moffitt_sample_id)
# Germ2 <- Germ2[(!grepl(uid, Germ2$avatar_id)),]
# Germ3 <-
#   readxl::read_xlsx(paste0(path,
#                            "/Raghu MM/Germline data/Germline_MM_Disease_Status_05052020_OUT .xlsx")) %>% 
#   select(c(avatar_id = "Avatar_id", "collectiondt", "WES_HUDSON_ALPHA", "Disease_Status"))
# Germ3 <- Germ3[(!grepl(uid, Germ3$avatar_id)),]
# Germ4 <-
#   readxl::read_xlsx(paste0(path,
#                            "/Raghu MM/Germline data/Moffitt_Germl_v0.4.5_Disease_Classification_OUT_07272020.xlsx")) %>% 
#   select(c("avatar_id", "SLID_germline", "Disease_Status"))
# Germ4 <- Germ4[(!grepl(uid, Germ4$avatar_id)),]
# 1.3.Load Sequencing data ---------------------------------------------------------------------------------------
WES_tumor <-
  readxl::read_xlsx(paste0(path, "/Raghu MM/Germline data/Moffitt_WES_v0.4.3_Disease_Classification_OUT01312020.xlsx")) %>% 
  select(c("avatar_id", "moffitt_sample_id", "collectiondt")) %>% 
  `colnames<-`(c("avatar_id", "moffitt_sample_id_tumor", "collectiondt_tumor"))

#---
Sequencing <-
  read.delim(paste0(path, "/Jamie/v0.4.3.MM.samples.WESdata01.31.20.txt")) %>% 
  select(c(
    "SLID_germline", 
    "SLID_tumor" , "moffitt_sample_id_tumor", 
    "moffitt_sample_id_germline",
    "BaitSet"))
# Sequencing$moffitt_sample_id_tumor == Sequencing$moffitt_sample_id # yes so remove one var
# Sequencing$subject == Sequencing$avatar_id # yes so remove one var
#---
Seq_WES_Raghu <- 
  readxl::read_xlsx(paste0(path, "/Raghu MM/Germline data/MM_Metadata_WES_V044.xlsx")) %>% 
  select(c(avatar_id = "subject", 
           "SLID_germline", "moffitt_sample_id_germline", "collectiondt_germline", 
           "SLID_tumor" , "moffitt_sample_id_tumor", "collectiondt_tumor", 
           "BaitSet"))
# Keep
# Seq_WES_Raghu$moffitt_sample_id_tumor == Seq_WES_Raghu$moffitt_sample_id # yes so rename and remove one var
# Seq_WES_Raghu$SLID_tumor == Seq_WES_Raghu$ClinicalSpecimenLinkage_WES # yes
# Seq_WES_Raghu$SLID_tumor == Seq_WES_Raghu$SLID # yes
# Seq_WES_Raghu$subject == Seq_WES_Raghu$ClinicalSpecimenLinkage_subject # yes
# Seq_WES_R$collectiondt_germline == Seq_WES_R$collectiondt_tumor # No -> That's good
#---
Sequencing2 <- # warning message due to a TRUE added in a num var by Raghu (he copy paste an extra patient)
  readxl::read_xlsx(paste0(path, "/Raghu MM/Germline data/MM_Metadata_WES_V0441.xlsx")) %>% 
  select(c(avatar_id = "subject",
           "SLID_germline", moffitt_sample_id_germline = "moffittSampleId_germline",
           "collectiondt_germline",
           "SLID_tumor" , moffitt_sample_id_tumor = "moffittSampleId_tumor", "collectiondt_tumor",
           BaitSet = "baitSet"))
#---
Seq_WES_Raghu2 <- 
  readxl::read_xlsx(paste0(path, "/Raghu MM/Germline data/MM_Metadata_WES_V045.xlsx")) %>% 
  select(c(avatar_id = "subject", 
           "SLID_germline", moffitt_sample_id_germline = "moffittSampleId_germline", "collectiondt_germline", 
           "SLID_tumor" , moffitt_sample_id_tumor = "moffittSampleId_tumor", "collectiondt_tumor", 
           "BaitSet"))

# 1.4.Load Clinical data------------------------------------------------------------------------------------------
# V1 and V2 verified in V4 format
ClinicalCap_V12 <-
  fs::path(
    "/Users/colinccm/Documents/GitHub/Gillis/no vpn data/CHIP in Avatar",
    "Raghu MM",
    "extracted Avatar V124 data and dict",
    "V1V2 verified by CIOX in V4 format"
  )
#---
Vitals_V12 <-
  readxl::read_xlsx((paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx")),
                    sheet = "Vitals") %>%
  select(c("avatar_id","vital_status","date_death", date_last_follow_up = "date_of_last_contact"))
# Remove IDs from new CIOX V1V2 in the initial V! and V2 version
uid_V <- paste(unique(Vitals_V12$avatar_id), collapse = '|')

Alc_Smo_V12 <-
  readxl::read_xlsx((paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx")),
                    sheet = "Comorbidities") %>%
  select(c("avatar_id","smoking_status", "alcohol_use"))
uid_A <- paste(unique(Alc_Smo_V12$avatar_id), collapse = '|')
#---
MM_history_V12 <-
  readxl::read_xlsx((paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx")),
                    sheet = "Myeloma_Disease_History") %>%
  select(c("avatar_id", "date_of_diagnosis"))
uid_MM <- paste(unique(MM_history_V12$avatar_id), collapse = '|')
#---
Treatment_V12 <-
  readxl::read_xlsx((paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx")),
                    sheet = "Treatment") %>%
  select(c("mrn", "avatar_id", "treatment_line_", "drug_start_date", "drug_name_", "drug_stop_date",
           "drug_name_other_")) %>% 
  unite(drug_name_, c(drug_name_,drug_name_other_), sep = ": ", na.rm = TRUE, remove = TRUE) # %>% 
  # rename(drug_other = "drug_name_other_")
uid_T <- paste(unique(Treatment_V12$avatar_id), collapse = '|')
#---
Progression_V12 <-
  readxl::read_xlsx((paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx")),
                    sheet = "Treatment_Outcomes") %>%
  select(c("avatar_id", progression_date = "initial_1_pd_date_1")) # For the non qc'd
Progr_V12 <-
  readxl::read_xlsx((paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx")),
                    sheet = "Treatment") %>%
  select(c("avatar_id", progression_date = "relapse_date", "QC")) # For the qc'd
uid_P <- paste(c(unique(Progression_V12$avatar_id), unique(Progr_V12$avatar_id)), collapse = '|')
#---
SCT_V12 <-
  readxl::read_xlsx((paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx")),
                    sheet = "SCT") %>%
  select(c("avatar_id","date_of_bmt")) %>%
  drop_na("date_of_bmt")
uid_S <- paste(unique(SCT_V12$avatar_id), collapse = '|')
#---
Radiation_V12 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx"),
                    sheet = "Radiation") %>%
  select(c("avatar_id", "rad_start_date", "rad_stop_date"))
uid_R <- paste(unique(Radiation_V12$avatar_id), collapse = '|')
#---
Labs_V12 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx"),
                    sheet = "Labs") %>%
  select(c("avatar_id", lab_date = "lab_initial_date_1"))
#---
Metastasis_V12 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx"),
                    sheet = "Metastatic_Disease") %>%
  select(c("avatar_id", metastasis_date = "mets_verify_date"))
#---
Staging_V12 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx"),
                    sheet = "Staging") %>%
  select(c("avatar_id", "date_staging_results"))
#---
Imaging_V12 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx"),
                    sheet = "Imaging") %>%
  select(c("avatar_id", "imaging_date"))
#---
Performance_V12 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx"),
                    sheet = "Performance") %>%
  select(c("avatar_id", "date_perf_status_dx"))
#---
TumorMarker_V12 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx"),
                    sheet = "Tumor_marker_flow") %>%
  select(c("avatar_id", "tumor_marker_date"))
#---
Biopsy_V12 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V12, "/Avatar_Legacy_V4_modif_09282020.xlsx"),
                    sheet = "Biopsy") %>%
  select(c("avatar_id", "biopsy_date"))

# V1 ----
ClinicalCap_V1 <-
  fs::path(
    "/Users/colinccm/Documents/GitHub/Gillis/no vpn data/CHIP in Avatar",
    "Raghu MM",
    "extracted Avatar V124 data and dict",
    "V1"
  )
#---
Vitals <-
  readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
                    sheet = "Vitals") %>%
  select(c("avatar_id","vital_status","date_death", "date_last_follow_up"))
Vitals <- Vitals[(!grepl(uid_V, Vitals$avatar_id)),]

Alc_Smo <-
  readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
                    sheet = "Vitals") %>%
  select(c("avatar_id", "smoking_status","current_smoker","alcohol_use"))
Alc_Smo <- Alc_Smo[(!grepl(uid_A, Alc_Smo$avatar_id)),]
#---
MM_history <-
  readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
                    sheet = "Myeloma_Disease_History") %>%
  select(c("avatar_id", "date_of_diagnosis", "disease_stage"))
MM_history <- MM_history[(!grepl(uid_MM, MM_history$avatar_id)),]
#---
# Comorbidities <-
#   readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
#                     sheet = "Comorbidities") #%>% 
#select(c("avatar_id","smoking_status", "alcohol_use"))
#---
Treatment <-
  readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
                    sheet = "Treatment") %>%
  select(c("avatar_id", drug_start_date = "regimen_start_date", drug_stop_date =  "regimen_end_date",
           "drug1_regimen", "drug2_regimen", "drug3_regimen", 
           "drug4_regimen", "drug5_regimen", "drug6_regimen", "drug7_regimen")) %>% 
  unite(drug_name_, drug1_regimen:drug7_regimen, sep = "; ", na.rm = TRUE, remove = TRUE)

Qcd_Treatment <-
  readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
                    sheet = "QC'd Treatment") %>%
  select(c("avatar_id", drug_start_date = "regimen_start_date", drug_stop_date = "regimen_end_date",
           drug_name_ = "treatment"))
Treatment <- Treatment[(!grepl(uid_T, Treatment$avatar_id)),]
Qcd_Treatment <- Qcd_Treatment[(!grepl(uid_T, Qcd_Treatment$avatar_id)),]
#---
Progression <-
  readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
                    sheet = "Treatment") %>%
  select(c("avatar_id", progression_date = "relapse_date"))
Progression <- Progression[(!grepl(uid_P, Progression$avatar_id)),]
#---
SCT <-
  readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
                    sheet = "SCT") %>%
  select(c("avatar_id","date_of_first_bmt", "date_of_second_bmt", "date_of_third_bmt"))
SCT <- SCT[(!grepl(uid_S, SCT$avatar_id)),]
#---
RadiationV1 <- readxl::read_xlsx(paste0(ClinicalCap_V1, "/Radiation_Version1_Patients.xlsx")) %>%
  select(c("Avatar_ID", "Radiation Start Date", "Radiation End Date")) %>% 
  `colnames<-`(c("avatar_id", "rad_start_date", "rad_stop_date"))
RadiationV1 <- RadiationV1[(!grepl(uid_R, RadiationV1$avatar_id)),]
#---
# Cytogenetic <-
#   readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
#                     sheet = "QC'd_Cytogenetics") %>%
#   select(c("avatar_id","date_bonemarrow_biopsy_results"))
#---
Biopsy <-
  readxl::read_xlsx((paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx")),
                    sheet = "Biopsy") %>%
  select(c("avatar_id", biopsy_date = "date_bonemarrow_biopsy_results"))
#---
Staging <- 
  readxl::read_xlsx(paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx"),
                    sheet = "Biopsy") %>%
  select(c("avatar_id", "date_staging_results"))
#---
Imaging <- 
  readxl::read_xlsx(paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx"),
                    sheet = "Biopsy") %>%
  select(c("avatar_id", imaging_date = "date_radiologicexam"))
#---
LabsV1 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V1, "/Avatar_MM_Clinical_Data_V1_modif_04292020.xlsx"),
                    sheet = "Biopsy") %>%
  select(c("avatar_id", "date_upep", "date_spep", "date_paraprotein_results",
           "date_flowcytometry_dna_"))
# V2 ----
ClinicalCap_V2 <-
  fs::path(
    "/Users/colinccm/Documents/GitHub/Gillis/no vpn data/CHIP in Avatar",
    "Raghu MM",
    "extracted Avatar V124 data and dict",
    "V2"
  )
#---
VitalsV2 <-
  readxl::read_xlsx((paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx")),
                    sheet = "Vitals") %>%
  select(c("avatar_id","vital_status","date_death"))
VitalsV2 <- VitalsV2[(!grepl(uid_V, VitalsV2$avatar_id)),]

Alc_Smo_V2 <-
  readxl::read_xlsx((paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx")),
                    sheet = "Vitals") %>%
  select(c("avatar_id", "smoking_status","alcohol_use")) # remove BMI
Alc_Smo_V2 <- Alc_Smo_V2[(!grepl(uid_A, Alc_Smo_V2$avatar_id)),]
#---
MM_historyV2 <-
  readxl::read_xlsx((paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx")),
                    sheet = "Myeloma_Disease_History") %>%
  select(c("avatar_id",  "date_of_diagnosis"))
MM_historyV2 <- MM_historyV2[(!grepl(uid_MM, MM_historyV2$avatar_id)),]
#---
TreatmentV2 <-
  readxl::read_xlsx((paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx")),
                    sheet = "Treatment") %>%
  select(c("avatar_id", "treatment_line_", "drug_start_date" , "drug_name_", "drug_stop_date",
           "drug_name_other")) %>% 
  unite(drug_name_, c(drug_name_,drug_name_other), sep = ": ", na.rm = TRUE, remove = TRUE)
Qcd_TreatmentV2 <-
  readxl::read_xlsx((paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx")),
                    sheet = "QC'd Treatment") %>%
  select(-relapse_date, -"...7")
TreatmentV2 <- TreatmentV2[(!grepl(uid_T, TreatmentV2$avatar_id)),]
Qcd_TreatmentV2 <- Qcd_TreatmentV2[(!grepl(uid_T, Qcd_TreatmentV2$avatar_id)),]
#---
ProgressionV2 <-
  readxl::read_xlsx((paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx")),
                    sheet = "Treatment") %>%
  select(c("avatar_id", progression_date = "relapse_date"))
ProgressionV2 <- ProgressionV2[(!grepl(uid_P, ProgressionV2$avatar_id)),]
#---
SCTV2 <-
  readxl::read_xlsx((paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx")),
                    sheet = "SCT") %>%
  select(c("avatar_id", "date_of_first_bmt", "date_of_second_bmt", "date_of_third_bmt"))
SCTV2 <- SCTV2[(!grepl(uid_S, SCTV2$avatar_id)),]
#---
RadiationV2 <- readxl::read_xlsx((paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx")),
                                 sheet = "Radiation") %>%
  select(c("avatar_id", "rad_start_date_v2", "rad_stop_date_v2")) %>% 
  `colnames<-`(c("avatar_id", "rad_start_date", "rad_stop_date"))
RadiationV2 <- RadiationV2[(!grepl(uid_R, RadiationV2$avatar_id)),]
#---
BiopsyV2 <-
  readxl::read_xlsx((paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx")),
                    sheet = "Biopsy") %>%
  select(c("avatar_id", biopsy_date = "date_bonemarrow_biopsy_results"))
#---
PerformanceV2 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx"),
                    sheet = "Performance") %>%
  select(c("avatar_id", date_perf_status_dx = "date_perf_status"))
#---
ImagingV2 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx"),
                    sheet = "Imaging") %>%
  select(c("avatar_id", imaging_date = "date_radiologicexam"))
#---
StagingV2 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx"),
                    sheet = "Staging") %>%
  select(c("avatar_id", "date_staging_results"))
#---
LabsV2 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V2, "/Avatar_MM_Clinical_Data_V2_modif_05042020.xlsx"),
                    sheet = "Biopsy") %>%
  select(c("avatar_id", "date_upep", "date_spep", "date_paraprotein_results"))
# V4 ----
ClinicalCap_V4 <-
  fs::path(
    "/Users/colinccm/Documents/GitHub/Gillis/no vpn data/CHIP in Avatar",
    "Raghu MM",
    "extracted Avatar V124 data and dict",
    "V4"
  )
#---
VitalsV4 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx")),
                    sheet = "Vitals") %>%
  select(c("avatar_id","vital_status","date_death", date_last_follow_up = "date_of_last_contact"))
#---
MM_historyV4 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx")),
                    sheet = "Myeloma_Disease_History") %>%
  select(c("avatar_id", "date_of_diagnosis"))
#---
# ComorbiditiesV4 <-
#   readxl::read_xlsx((paste0(ClinicalCap_V4,"/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx")),
#                     sheet = "Comorbidities") %>%
#   select(c("avatar_id", ))
Alc_SmoV4 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx")),
                    sheet = "Comorbidities") %>%
  select(c("avatar_id","smoking_status", "alcohol_use"))
#---
TreatmentV4 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx")),
                    sheet = "Treatment") %>%
  select(c("avatar_id", "treatment_line_", "drug_start_date", "drug_name_", "drug_stop_date",
           "drug_name_other")) %>% 
  unite(drug_name_, c(drug_name_,drug_name_other), sep = ": ", na.rm = TRUE, remove = TRUE)# %>% 
# rename(drug_other = "drug_name_other")
#---
Progression_V4 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx")),
                    sheet = "Treatment_Outcomes") %>%
  select(c("avatar_id", progression_date = "initial_1_pd_date_1"))
#---
SCTV4 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx")),
                    sheet = "SCT") %>%
  select(c("avatar_id", "date_of_bmt")) %>%  # can be different than first if duplicated from v1 or v2
  drop_na("date_of_bmt")
# SCTV4 <- dcast(setDT(SCTV4), avatar_id ~ rowid(avatar_id), value.var = c("date_of_bmt")) %>% 
#   rename("date_of_first_bmt" = "1", "date_of_second_bmt" = "2")
#---
RadiationV4 <- 
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx")),
                    sheet = "Radiation") %>%
  select(c("avatar_id", "rad_start_date", "rad_stop_date"))
#---
LabsV4 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx"),
                    sheet = "Labs") %>%
  select(c("avatar_id", lab_date = "lab_initial_date_1"))
#---
MetastasisV4 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx"),
                    sheet = "Metastatic_Disease") %>%
  select(c("avatar_id", metastasis_date = "mets_verify_date"))
#---
StagingV4 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx"),
                    sheet = "Staging") %>%
  select(c("avatar_id", "date_staging_results"))
#---
ImagingV4 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx"),
                    sheet = "Imaging") %>%
  select(c("avatar_id", "imaging_date"))
#---
PerformanceV4 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx"),
                    sheet = "Performance") %>%
  select(c("avatar_id", "date_perf_status_dx"))
#---
TumorMarkerV4 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx"),
                    sheet = "Tumor_marker_flow") %>%
  select(c("avatar_id", "tumor_marker_date"))
#---
BiopsyV4 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_modif_04272020.xlsx"),
                    sheet = "Biopsy") %>%
  select(c("avatar_id", biopsy_date = "date_bonemarrow_biopsy_results"))

# V4.august2020 ----
#
VitalsV4.1 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx")),
                    sheet = "Vitals") %>%
  select(c("avatar_id","vital_status","date_death", date_last_follow_up = "date_of_last_contact"))
#
MM_historyV4.1 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx")),
                    sheet = "Myeloma_Disease_History") %>%
  select(c("avatar_id", "date_of_diagnosis"))
#
Alc_SmoV4.1 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx")),
                    sheet = "Comorbidities") %>%
  select(c("avatar_id","smoking_status", "alcohol_use"))
#
TreatmentV4.1 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx")),
                    sheet = "Treatment") %>%
  select(c("avatar_id", "treatment_line_", "drug_start_date", "drug_name_", "drug_stop_date",
           "drug_name_other_")) %>% 
  unite(drug_name_, c(drug_name_,drug_name_other_), sep = ": ", na.rm = TRUE, remove = TRUE)
#
Progression_V4.1 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx")),
                    sheet = "Treatment_Outcomes") %>%
  select(c("avatar_id", progression_date = "initial_1_pd_date_1"))
#
SCTV4.1 <-
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx")),
                    sheet = "SCT") %>%
  select(c("avatar_id", "date_of_bmt")) %>%  # can be different than first if duplicated from v1 or v2
  drop_na("date_of_bmt")
# SCTV4.1 <- dcast(setDT(SCTV4), avatar_id ~ rowid(avatar_id), value.var = c("date_of_bmt")) %>% 
#   rename("date_of_first_bmt" = "1", "date_of_second_bmt" = "2")
#
RadiationV4.1 <- 
  readxl::read_xlsx((paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx")),
                    sheet = "Radiation") %>%
  select(c("avatar_id", "rad_start_date", "rad_stop_date"))
#
LabsV4.1 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx"),
                    sheet = "Labs") %>%
  select(c("avatar_id", lab_date = "lab_initial_date_1"))
#
MetastasisV4.1 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx"),
                    sheet = "Metastatic_Disease") %>%
  select(c("avatar_id", metastasis_date = "mets_verify_date"))
#
StagingV4.1 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx"),
                    sheet = "Staging") %>%
  select(c("avatar_id", "date_staging_results"))
#
ImagingV4.1 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx"),
                    sheet = "Imaging") %>%
  select(c("avatar_id", "imaging_date"))
#
PerformanceV4.1 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx"),
                    sheet = "Performance") %>%
  select(c("avatar_id", "date_perf_status_dx"))
#
TumorMarkerV4.1 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx"),
                    sheet = "Tumor_marker_flow") %>%
  select(c("avatar_id", "tumor_marker_date"))
#
BiopsyV4.1 <- 
  readxl::read_xlsx(paste0(ClinicalCap_V4, "/Avatar_MM_Clinical_Data_V4_OUT_08032020 .xlsx"),
                    sheet = "Biopsy") %>%
  select(c("avatar_id", biopsy_date = "date_bonemarrow_biopsy_results"))
#
OS_data <- readxl::read_xlsx(paste0(path, "/Raghu MM/Overall Survival/HRI_Last_Followupdata.xlsx")) %>% 
  select(avatar_id = "germline_patient_data_avatar_id", "final_vitals", "Vital_Status_Date") %>% 
  distinct()
#
Staging_ISS <- readxl::read_xlsx(paste0(path, "/Raghu MM/Staging_09142020.xlsx")) %>% 
  select("avatar_id", "collectiondt_germline", "Labs_Result_Date", "Final_Albumin", "Final_Beta2", "Final_LDH", "ISS") %>% 
  mutate(ISS = str_replace(ISS, pattern = "NA", replacement = NA_character_)) %>% 
  arrange(collectiondt_germline) %>% 
  distinct(avatar_id, .keep_all = TRUE) # remove the duplicate of patient 180
#
CHIP_status <- read_csv(paste0(path, "/Nancy's working files/CHcalls_12.10.20.csv")) %>% 
  # mutate(CH_status = ifelse(CH_status == "CH", "CHIP", "No CHIP")) %>% 
  mutate(patient_germline_id = str_remove(patient_germline_id, "_normal"))
#
IMIDS_maintenance <- readxl::read_xlsx(paste0(path, "/Raghu MM/MM_Maintainance_Regimen.xlsx"))
#
migration_patients <- readxl::read_xlsx(paste0(path, "/Raghu MM/Avatar List For Migration.xlsx"))

path <- fs::path("","Volumes","Gillis_Research","Christelle Colin-Leitzinger", "CHIP in Avatar")
```

In total, we have 4 versions of the clinical data.  
V1 has `r MM_history %>% distinct(avatar_id) %>% NROW()` patients and was sent by Raghu on Feb, 07th 2020.  
V2 has `r MM_historyV2 %>% distinct(avatar_id) %>% NROW()` patients and was sent by Raghu on Feb, 10th 2020.  
V4 has `r MM_historyV4 %>% distinct(avatar_id) %>% NROW()` patients and was sent by Raghu on April, 21st 2020.  
A V4.1 data added `r MM_historyV4.1 %>% distinct(avatar_id) %>% NROW()` patients to V4 on August, 03rd 2020.  
V12 has `r MM_history_V12 %>% distinct(avatar_id) %>% NROW()` patients and was sent by Raghu on Sept, 04th 2020. This data overlap V1 and V2 patients with some patients data which were updated by CIOX and reformatted to a V4 format, and patients for who the data were only reformatted to a V4 format but not verified.  
We learned in Feb 2021 that CIOX contract was stopped.  
Radiation data are not present in V1 so Raghu sent me another file with them pulled from HRI on March 14th 2020.  
The ISS staging was calculated by Raghu and sent on Sept 14th 2020. It doesn't use cytogenetics features. The ISS is calculated on all patients MM, Smoldering and Mgus, a later cleaning only kept the ISS for the patients who are MM ("Pre Treatment Newly Diagnosed Multiple Myeloma","Post Treatment Newly Diagnosed Multiple Myeloma","Early Relapse Multiple Myeloma","Late Relapse Multiple Myeloma") at the time of germline collection.  
The IMIDS maintenance file was sent by Raghu on Dec 29th 2020 that he queried from raw datasets of Avatar legacy v4 MM cases which matched to our patient list. There are about 135 cases (from v045) that are awaiting qc of migration(clinical data), follow up data. I have not used these cases for querying and about 90 cases of these patients list are v1 cases where forms were not designed to capture maintenance regimen data.  
Comments: Raghu said that "Original data (regimen line variable) comes as numbers with different values in V4 form. There is some missing data and some values are not very relevant to Myeloma since V4 forms are not disease specific, so Dr. Shain recommended me to reformat the variable."  

## Germline
```{r}
Germline <- Germline %>% 
  distinct()

Sequencing <-
  full_join(
    Sequencing,
    WES_tumor,
    by = "moffitt_sample_id_tumor")
# Bind Sequencing
Seq_WES <- bind_rows(Seq_WES_Raghu, Sequencing2, Seq_WES_Raghu2, Sequencing, .id = "vers") %>% 
  arrange(collectiondt_germline) %>% 
  distinct(SLID_tumor, moffitt_sample_id_tumor, SLID_germline, .keep_all = TRUE)

Seq_WES <- Seq_WES[order(Seq_WES$collectiondt_tumor), ]
# pivot wider
WES_seq <-
  dcast(setDT(Seq_WES), avatar_id+SLID_germline+moffitt_sample_id_germline+collectiondt_germline ~ rowid(avatar_id),
        value.var = c(
          "SLID_tumor",
          "moffitt_sample_id_tumor",
          "collectiondt_tumor",
          "BaitSet"
        )
  )

# Merge all
Germline <- left_join(WES_seq, Germline, by = "avatar_id") %>% 
  filter(SLID_germline.x == SLID_germline.y | is.na(SLID_germline.x == SLID_germline.y)) %>% 
  rename(SLID_germline = "SLID_germline.x", collectiondt_germline = "collectiondt_germline.x") %>% 
  # distinct(avatar_id, SLID_germline, .keep_all = TRUE) %>% 
  mutate(collectiondt_germline = coalesce(collectiondt_germline, collectiondt_germline.y)) %>% 
  select(-SLID_germline.y, -collectiondt_germline.y)
```

## Demographics
```{r}

```


## MM history cleaning
```{r}
mm_history <- bind_rows(MM_history_V12, MM_history, MM_historyV2, MM_historyV4, MM_historyV4.1, .id = "versionMM") %>%
  drop_na("date_of_diagnosis") %>%
  distinct(avatar_id, date_of_diagnosis, .keep_all = TRUE) %>% 
  # Create date of diagnosis closest to germline date
  left_join(., Germline %>% distinct(avatar_id, .keep_all = TRUE) %>% select("avatar_id", "collectiondt_germline"), # For only 1 date of Dx when multiple germline collection
            by = "avatar_id") %>% 
  mutate(interval = (interval(start= .$collectiondt_germline, end= .$date_of_diagnosis)/duration(n=1, unit="days"))) %>% 
  mutate(interval1 = if_else(interval>100, NA_real_, interval)) %>% 
  mutate(interval1 = abs(interval1)) %>% 
  arrange(interval1) %>% 
  group_by(avatar_id) %>% mutate(id = 1:n()) %>% ungroup() %>%
  mutate(Dx_date_closest_germline = if_else(id == 1, date_of_diagnosis, NA_POSIXct_)) %>% 
  distinct(avatar_id, date_of_diagnosis, .keep_all = TRUE) %>% 
  group_by(avatar_id) %>% fill(Dx_date_closest_germline, .direction = "updown") %>% 
  arrange(avatar_id, date_of_diagnosis) %>% 
  ungroup()

MM_history <- dcast(setDT(mm_history), avatar_id+Dx_date_closest_germline ~ rowid(avatar_id), 
                    value.var = c("date_of_diagnosis", "disease_stage")) %>% 
  mutate(date_of_MMonly_diagnosis = case_when(
    disease_stage_1 == "active"          ~ date_of_diagnosis_1,
    disease_stage_2 == "active"          ~ date_of_diagnosis_2,
    disease_stage_3 == "active"          ~ date_of_diagnosis_3,
    disease_stage_4 == "active"          ~ date_of_diagnosis_4,
  )) %>% 
  mutate(date_of_MMSMMGUSdiagnosis = coalesce(date_of_MMonly_diagnosis, date_of_diagnosis_1)) %>% 
  select(c("avatar_id", "Dx_date_closest_germline", "date_of_MMonly_diagnosis", "date_of_MMSMMGUSdiagnosis", everything()))
```
All MM history are bind together and duplicated rows are removed.
The date of diagnosis are multiple for each patients. It could be the first time they got a Mgus, Smoldering or MM diagnosis. It can also be when the patient become MM or when he relapse. The `disease_stage` associated with the `date_of_diagnosis` tells us if the patient is within this 4 conditions.  
`r table(MM_history$disease_stage_1)`
By binding with the `collectiondt_germline` I was able to calculate an interval between the later and the multiple `date_of_diagnosis`.
We choose to take the date closest to germline (with closest absolut interval as long as it is not nore than 100 days after the germline) as the `Dx_date_closest_germline`. <span style='color: red;'>The problem can be when the disease stage is not recorded.</span>  
I also created a `date_of_MMonly_diagnosis` variable which take the first diagnosis date where the disease status is active (MM diagnosis). Comment: it doesn't take the relapse diagnosis...<span style='color: red;'>Should it?</span>  
`date_of_MMSMMGUSdiagnosis` takes `date_of_MMonly_diagnosis` and NA are filled with the very first diagnosis date of non-active disease in the data.
```{r}
mm_history
MM_history
```








