---
title: "MM Avatar cleaning"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cerulean #darkly
    highlight: pygments
    df_print: paged
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   <!-- margin-bottom: 100px; -->
}

table {
    margin-top: 25px;
    <!-- margin-bottom: 100px !important; -->
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")

```

<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">Workflow through the MM Avatar data cleaning</span>

</div>
<br>

```{r library, include=FALSE}
library(tidyverse)
library(data.table)
library(lubridate)
```

***
# I.Clinical data

```{r load}

```

In total, we have 4 version of the clinical data.
V1 has `r ` patients and was sent by Raghu on Feb, 07 2020.
V2 has `r ` patients and was sent by Raghu on Feb, 10 2020.
V4 has `r ` patients and was sent by Raghu on April, 21 2020. 
A V4.1 data added `r ` patients to V4 on August, 03 2020.
V12 has `r ` patients and was sent by Raghu on Sept, 04 2020. This data overlap V1 and V2 patients with some patients data which were updated by CIOX and reformatted to a V4 format, and patients for who the data were only reformatted to a V4 format but not verified.
We learned in Feb 2021 that CIOX contract was stopped.













