---
title: "Disparities Prop Hazard Assumption"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: no
    theme: cerulean
    highlight: pygments
editor_options:
  chunk_output_type: console
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   margin-bottom: 25px;
}

table {
    margin-top: 25px;
    <!-- margin-bottom: 100px !important; -->
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 12, data_row.padding = gt::px(1))")

```


<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">Proportional Assumption</span>

</div>
<br>

```{r library, include=FALSE}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(viridis)
library(lubridate)
library(gtsummary)
library(survival)
library(survminer)
```

```{r load}
# Global_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Global_data.rds")
germline_patient_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/germline_patient_data.rds")
path <- ("/Users/colinccm/Documents/GitHub/Gillis/no vpn data/CHIP in Avatar")

germline_patient_data <- germline_patient_data %>% 
  rename(Sex = Gender) %>% 
  filter(is_patient_MM == "Yes" & is_MMDx_close_to_blood == "Yes") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  filter(raceeth != "Others") %>% 
  mutate(raceeth = factor(raceeth, levels=c("White Non-Hispanic", "Hispanic", "Black"))) %>%
  mutate(ISS_at_MMdx = replace_na(ISS_at_MMdx, "Unknown")) %>%
  mutate(regimen_categoryVCD_1 = factor(regimen_categoryVCD_1, levels = c(
    "PI + IMIDs", "PI", "IMIDs", "Others"
  )))
```
<br>
<br>

***

# Introduction

The ORIEN Avatar^TM Research Program focuses on patients with advanced primary or metastatic disease, those with limited treatment options, as well as patients who are likely to develop progressive disease, and will capitalize on the disease-area expertise of ORIEN's research scientists.   
It generates massive amounts of genetic and clinical information on patients consenting to the Total Cancer Care® (TCC) Protocol.   
ORIEN Avatar tracks patient clinical progressions over their lifetime to provide the richest data source for targeted cancer therapy advancement.  ORIEN Avatar data solution is optimized for oncology research by providing integrated longitudinal clinical data and whole exome, transcriptome and germline sequencing around consented oncology patients.   
We are focusing on patients with multiple myeloma disease.
We will focus on patient with active MM who had WES analyzed on their blood and CH were analyzed.

<br>

***

# I. MM Avatar data description


```{r}
germline_patient_data <- germline_patient_data %>% 
  mutate(HCT_comparison = case_when(
    is.na(date_of_bmt_1)              ~ "No HCT",
    time_to_hct < 314                 ~ "Early HCT",
    time_to_hct >= 314                ~ "Late HCT"
  )) %>% 
  mutate(HCT_comparison = factor(HCT_comparison, levels = c("No HCT", "Early HCT", "Late HCT")))

contamination_tumor_ids <- paste("A025767", "A025771", "A007368", "A007322", "A000111", sep = "|")
```


***

# II. Age Distribution in Disparities


***

# III. RaceEthnicity

## Survivals

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
raceeth_surv <- germline_patient_data

mysurv <- Surv(time = raceeth_surv$month_at_os, event = raceeth_surv$os_event)
myplot <- survfit(mysurv~raceeth, data = raceeth_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = raceeth_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           palette = c("blue", "darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3))

# tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = raceeth_surv$month_at_os, 
  #                            event = raceeth_surv$os_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = raceeth_surv$month_at_os, 
             event = raceeth_surv$os_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever, data =  raceeth_surv) %>%
  cox.zph()
# eth_cox_os <- 
  # tbl_merge(list(# tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
# eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = raceeth_surv$month_at_progression_drug, event = raceeth_surv$drug_progression_event)
myplot <- survfit(mysurv~raceeth, data = raceeth_surv)
ggsurvplot(myplot, data = raceeth_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           palette = c("blue", "darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3))

# tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = raceeth_surv$month_at_progression_drug, 
  #                            event = raceeth_surv$drug_progression_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = raceeth_surv$month_at_progression_drug, 
             event = raceeth_surv$drug_progression_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever), data =  raceeth_surv) %>%
  cox.zph()
# eth_cox_pfs <- 
  # tbl_merge(list(# tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
# eth_cox_pfs
```
</div>
</div>

<br>
<br>


***

# IV. CH

From here, remove 5 patients with tumor contamination

```{r}
patients_clean_tumor <- germline_patient_data %>% 
  filter(!str_detect(avatar_id, contamination_tumor_ids))
```

```{r}
ch_char <- patients_clean_tumor %>% 
  mutate(Whole = "MM patients at time of blood with WES") %>% 
  select(Whole, CH_status) %>%
  tbl_summary(by = Whole, 
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels()
ch_char
```
<br>
<br>

***

# IV. CH & RaceEthnicity
### PFS
```{r, fig.height = 7}
mysurv <- Surv(time = patients_clean_tumor$month_at_progression_drug, event = patients_clean_tumor$drug_progression_event)
myplot <- survfit(mysurv~raceeth + CH_status, data = patients_clean_tumor)
ggsurvplot(myplot, data = patients_clean_tumor,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "raceeth",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 3), linetype = guide_legend(nrow = 2))
```

```{r}
# tbl1 <- patients_clean_tumor %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, CH_status) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = patients_clean_tumor$month_at_progression_drug, 
  #                            event = patients_clean_tumor$drug_progression_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = patients_clean_tumor$month_at_progression_drug, 
             event = patients_clean_tumor$drug_progression_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + CH_status, data =  patients_clean_tumor) %>%
  cox.zph()
# tbl_merge(list(# tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

### OS
```{r, fig.height = 7}
mysurv <- Surv(time = patients_clean_tumor$month_at_os, event = patients_clean_tumor$os_event)
myplot <- survfit(mysurv~raceeth + CH_status, data = patients_clean_tumor)
ggsurvplot(myplot, data = patients_clean_tumor,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "raceeth",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 3), linetype = guide_legend(nrow = 2))
```

```{r}
# tbl1 <- patients_clean_tumor %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, CH_status) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = patients_clean_tumor$month_at_os, 
  #                            event = patients_clean_tumor$os_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = patients_clean_tumor$month_at_os, 
                   event = patients_clean_tumor$os_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + CH_status, data =  patients_clean_tumor) %>%
  cox.zph()
# tbl_merge(list(# tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
<br>

***

<br>
<br>

## WNH
<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
nhis_surv <- patients_clean_tumor %>% 
  filter(raceeth == "White Non-Hispanic")

mysurv <- Surv(time = nhis_surv$month_at_os, event = nhis_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = nhis_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = nhis_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           # palette = c("darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
)

# tbl1 <- nhis_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, Drugs_ever, HCT_ever) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = nhis_surv$month_at_os, 
  #                            event = nhis_surv$os_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = nhis_surv$month_at_os, 
             event = nhis_surv$os_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + Drugs_ever + HCT_ever, data =  nhis_surv) %>%
  cox.zph()
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = nhis_surv$month_at_progression_drug, event = nhis_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = nhis_surv)
ggsurvplot(myplot, data = nhis_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           palette = c("darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

# tbl1 <- nhis_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = nhis_surv$month_at_progression_drug, 
  #                            event = nhis_surv$drug_progression_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = nhis_surv$month_at_progression_drug, 
             event = nhis_surv$drug_progression_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever, data =  nhis_surv) %>%
  cox.zph()

```
</div>
</div>

<br>
<br>

## Bl
<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
bl_surv <- patients_clean_tumor %>% 
  filter(raceeth == "Black")

mysurv <- Surv(time = bl_surv$month_at_os, event = bl_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = bl_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = bl_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           # palette = c("darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
)

# tbl1 <- bl_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, Drugs_ever, HCT_ever) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = bl_surv$month_at_os, 
  #                            event = bl_surv$os_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = bl_surv$month_at_os, 
             event = bl_surv$os_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + Drugs_ever + HCT_ever, data =  bl_surv) %>%
  cox.zph()

```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = bl_surv$month_at_progression_drug, event = bl_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = bl_surv)
ggsurvplot(myplot, data = bl_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           palette = c("darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

# tbl1 <- bl_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = bl_surv$month_at_progression_drug, 
  #                            event = bl_surv$drug_progression_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = bl_surv$month_at_progression_drug, 
             event = bl_surv$drug_progression_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever, data =  bl_surv) %>%
  cox.zph()

```
</div>
</div>

<br>
<br>

## H

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
his_surv <- patients_clean_tumor %>% 
  filter(raceeth == "Hispanic")

mysurv <- Surv(time = his_surv$month_at_os, event = his_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = his_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = his_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           # palette = c("darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
)

# tbl1 <- his_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, Drugs_ever, HCT_ever) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = his_surv$month_at_os, 
  #                            event = his_surv$os_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = his_surv$month_at_os, 
             event = his_surv$os_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + Drugs_ever + HCT_ever, data =  his_surv) %>%
  cox.zph()

```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = his_surv$month_at_progression_drug, event = his_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = his_surv)
ggsurvplot(myplot, data = his_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           palette = c("darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

# tbl1 <- his_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever) %>% 
  # tbl_uvregression(method = survival::coxph, 
  #                  y = (Surv(time = his_surv$month_at_progression_drug, 
  #                            event = his_surv$drug_progression_event)),
  #                  exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  # bold_labels() %>% italicize_levels()
coxph(Surv(time = his_surv$month_at_progression_drug, 
             event = his_surv$drug_progression_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever, data =  his_surv) %>%
  cox.zph()

```
</div>
</div>

<br>


<br>

<!-- I thought it could be a year effect compare to Taiga said (maybe earlier it wasn't a difference like he said but now that Hispanic may have more access to care we can see the difference) but no. -->

```{r}
# mysurv <- Surv(time = ethnicity_surv$month_at_progression_drug, event = ethnicity_surv$drug_progression_event)
# myplot <- survfit(mysurv~Ethnicity1+diagnosis_MM_year, data = ethnicity_surv)
# ggsurvplot_facet(myplot, data = ethnicity_surv, facet.by = "diagnosis_MM_year",
#            title = "PFS",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(10, "bold", "black"),
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months",
#            palette = c("darkred", "darkgreen"),
#            conf.int = FALSE
# )
```

***

# Not shown

## CH

### Survivals

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = patients_clean_tumor$month_at_os, event = patients_clean_tumor$os_event)
myplot <- survfit(mysurv~CH_status, data = patients_clean_tumor)
# jpeg(paste0(path, "/Figures/Survivals/CHIP/OS by CH.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = patients_clean_tumor,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "", 
           # # legend.labs = c("CH", "No CH"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.17,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black"))
)
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = patients_clean_tumor$month_at_progression_drug, event = patients_clean_tumor$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = patients_clean_tumor)
ggsurvplot(myplot, data = patients_clean_tumor,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "", 
           # # legend.labs = c("CH", "No CH"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.17,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black"))
)
```
</div>
</div>

***
