---
title: "Disparities"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: no
    theme: cerulean
    highlight: pygments
editor_options:
  chunk_output_type: console
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   margin-bottom: 25px;
}

table {
    margin-top: 25px;
    <!-- margin-bottom: 100px !important; -->
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 12, data_row.padding = gt::px(1))")

```


<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">Manuscript</span>

</div>
<br>

```{r library, include=FALSE}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(viridis)
library(lubridate)
library(gtsummary)
library(survival)
library(survminer)
```

```{r load}
# Global_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Global_data.rds")
germline_patient_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/germline_patient_data.rds")
path <- ("/Users/colinccm/Documents/GitHub/Gillis/no vpn data/CHIP in Avatar")

germline_patient_data <- germline_patient_data %>% 
  filter(is_patient_MM == "Yes" & is_MMDx_close_to_blood == "Yes") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(CH_status = str_replace(CH_status, "No_CH", "No CH"))
```
<br>
<br>

***

# Introduction

The ORIEN Avatar^TM Research Program focuses on patients with advanced primary or metastatic disease, those with limited treatment options, as well as patients who are likely to develop progressive disease, and will capitalize on the disease-area expertise of ORIEN's research scientists.   
It generates massive amounts of genetic and clinical information on patients consenting to the Total Cancer Care® (TCC) Protocol.   
ORIEN Avatar tracks patient clinical progressions over their lifetime to provide the richest data source for targeted cancer therapy advancement.  ORIEN Avatar data solution is optimized for oncology research by providing integrated longitudinal clinical data and whole exome, transcriptome and germline sequencing around consented oncology patients.   
We are focusing on patients with multiple myeloma disease.
We will focus on patient with active MM who had WES analyzed on their blood and CH were analyzed.

<br>

***

# I. MM Avatar data description

## 1.Demographics of MM patients with WES
FTI, I left the "Others" in raceeth.

<div class = "row">
<div class = "col-md-5">
```{r}
germ_data <- germline_patient_data %>% 
  mutate(Whole = "MM patients with WES") %>% 
  select(Age_at_MMonly_diagnosis, Gender, raceeth, Race, Race1, Ethnicity, Ethnicity1, Whole, ISS, ISS_grp, ISS_at_MMdx, ISSdx_grp) %>%
  tbl_summary(by = Whole, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric", ISS_grp ~ "alphanumeric", ISS_at_MMdx ~ "alphanumeric", ISSdx_grp ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels()
germ_data
```
</div>

<div class = "col-md-7">
```{r}
p <- qplot(x =Age_at_MMonly_diagnosis, data=subset(germline_patient_data,!is.na(Age_at_MMonly_diagnosis)), fill=..count.., geom="histogram") 
p + scale_fill_viridis_c(
  alpha = 1,
  begin = 0,
  end = 1,
  direction = 1,
  option = "D",
  values = NULL,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
) +
  theme_minimal(base_size = 16) +
  labs(x="Age at Diagnosis", y="Number of Patient", title="Repartition of Age at Diagnosis in MM Avatar")
```
</div>
</div>

<br>

From here, I removed the "Others" in raceeth.

```{r}
germline_patient_data <- germline_patient_data %>% 
  filter(raceeth != "Others") %>% 
  mutate(raceeth = factor(raceeth, levels=c("White Non-Hispanic", "Hispanic", "Black")))

```

<div class = "row">
<div class = "col-md-5">
```{r}
germ_data <- germline_patient_data %>% 
  mutate(Whole = "MM patients with WES") %>% 
  select(Age_at_MMonly_diagnosis, Gender, raceeth, Race, Race1, Ethnicity, Ethnicity1, Whole, ISS, ISS_grp, ISS_at_MMdx, ISSdx_grp) %>%
  tbl_summary(by = Whole, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric", ISS_grp ~ "alphanumeric", ISS_at_MMdx ~ "alphanumeric", ISSdx_grp ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels()
germ_data
```
</div>

<div class = "col-md-7">
```{r}
p <- qplot(x =Age_at_MMonly_diagnosis, data=subset(germline_patient_data,!is.na(Age_at_MMonly_diagnosis)), fill=..count.., geom="histogram") 
p + scale_fill_viridis_c(
  alpha = 1,
  begin = 0,
  end = 1,
  direction = 1,
  option = "D",
  values = NULL,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
) +
  theme_minimal(base_size = 16) +
  labs(x="Age at Diagnosis", y="Number of Patient", title="Repartition of Age at Diagnosis in MM Avatar")
```
</div>
</div>

<br>

Patients represented in MM Avatar data are characterized by a median age at diagnosis of `r inline_text(germ_data, variable = Age_at_MMonly_diagnosis, column = "MM patients with WES", pattern = "{median} years old")`. We saw slightly more male `r inline_text(germ_data, variable = Gender, level = "Male", column = "MM patients with WES")`, a majority of White Non-Hispanic `r inline_text(germ_data, variable = Race1, level = "White", column = "MM patients with WES")` and Non-Hispanic `r inline_text(germ_data, variable = Ethnicity1, level = "Non-Hispanic", column = "MM patients with WES")`.
A higher proportion of patients have a lower clinical stage at diagnosis with `r inline_text(germ_data, variable = ISS_grp, level = "I", column = "MM patients with WES")` patients with ISS I and `r inline_text(germ_data, variable = ISS_grp, level = "II-III", column = "MM patients with WES")` patients with ISS II and III combined.


## 2. CH

```{r}
ch_char <- germline_patient_data %>% 
  mutate(Whole = "MM patients at time of blood with WES") %>% 
  select(Whole, CH_status) %>%
  tbl_summary(by = Whole, 
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels()
ch_char
```

Whole Exome Sequencing data identified CH mutations in `r inline_text(ch_char, variable = CH_status, level = "CH", column = "MM patients at time of blood with WES")`.
+ Add bioinfo results (oncoplot?).  
Type of genes affected...


```{r}
# ch_samples <- germline_patient_data %>%
#   select(CH_status) %>% 
#   tbl_summary(sort = list(everything() ~ "frequency")) %>% 
#   bold_labels()
# ch_samples

ch_demo <- germline_patient_data %>%
  select(Age_at_MMonly_diagnosis, 
         Gender, raceeth, Race1, Ethnicity1, ISS, ISS_grp, ISS_at_MMdx, ISSdx_grp, 
         CH_status) %>% 
  tbl_summary(by = CH_status, 
              label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis", CH_status ~ "CH status"),
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric", ISS_grp ~ "alphanumeric", ISS_at_MMdx ~ "alphanumeric", ISSdx_grp ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)
ch_demo
```
<br>

A significant difference in the age at diagnosis was observed between CH and No CH patients with an older age for CH patients (`r inline_text(ch_demo, variable = Age_at_MMonly_diagnosis, column = "CH")` vs `r inline_text(ch_demo, variable = Age_at_MMonly_diagnosis, column = "No CH")`, `r inline_text(ch_demo, variable = Age_at_MMonly_diagnosis, column = "p.value")`). While Gender and race/ethnicity are similarly distributed, we noticed a higher proportion of patients with more severe clinical stage (II-III) for CH patients (`r inline_text(ch_demo, variable = ISS_grp, column = "p.value")`).

<br>

## 3. Disease Status (Just for us?)

FYI  
These patients are now all MM but at time of blood collection distribution of disease status is as below.  
Few discrepancies I can remove...
```{r}
germline_patient_data %>%
  mutate(Whole = "MM patients with WES") %>% 
  select(Disease_Status_germline, Whole) %>% 
  tbl_summary(by = Whole, 
              label = list(Disease_Status_germline ~ "Disease Status at Blood Collection")) %>% 
  bold_labels() %>% as_gt()  %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

# misclassified <- germline_patient_data %>%
#   select(avatar_id,
#          Disease_Status_germline, collectiondt_germline, date_of_MMonly_diagnosis, is_patient_MM,
#          is_MMDx_close_to_blood, smoldering_status, 
#          date_of_diagnosis_1, disease_stage_1, 
#          date_of_diagnosis_2, disease_stage_2,
#          date_of_diagnosis_3, disease_stage_3,
#          date_of_diagnosis_4, disease_stage_4) %>% 
#   filter(str_detect(Disease_Status_germline, "Soli|Smol|Mgus|Amyl|Myelofib")) %>% 
#   arrange(Disease_Status_germline)

germline_patient_data %>%
  group_by(Disease_Status_germline) %>% count(Disease_Status_germline) %>% 
  mutate(Disease_Status_germline = fct_reorder(as.character(Disease_Status_germline), desc(n))) %>%
  arrange(n) %>% 
  ggplot(aes(x="", y=n, fill=Disease_Status_germline)) +
  geom_bar(stat="identity", width=1) +
  scale_fill_manual("", values = c("hotpink", "orchid1", "mediumorchid1", "mediumpurple1", "lemonchiffon", "peachpuff", 
                                                               "lightcyan", "lightblue", "cyan", "steelblue1", "blue", "darkblue")) +
  theme_void(base_size = 16) +
  theme(plot.title = element_text(hjust = -0.15))+
  coord_polar("y", start=0, direction=-1) +
  labs(x=NULL, y=NULL, title="Disease Status at Blood Collection in MM Avatar Data")
```


# II. Age Distribution in Disparities

## Raceeth
```{r, fig.height = 10}
p <- ggplot(germline_patient_data, 
            aes(x=raceeth, y=Age_at_MMonly_diagnosis), fill=raceeth) + 
  geom_boxplot(color = c("darkred", "darkgreen", "blue")) + 
  theme_minimal(base_size = 23) +
  labs(x=NULL, y="Age at Diagnosis", title="Age of MM Diagnosis Distribution")
p + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means()
```

## Separated race / ethn
<div class = "row">
<div class = "col-md-5">
```{r, fig.height = 10}
p <- ggplot(germline_patient_data, 
            aes(x=Ethnicity1, y=Age_at_MMonly_diagnosis), fill=Ethnicity1) + 
  geom_boxplot(color = c("darkred", "darkgreen")) + 
  theme_minimal(base_size = 23) +
  labs(x=NULL, y="Age at Diagnosis", title="Age of MM Diagnosis Distribution")
p + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means()
```
</div>

<div class = "col-md-7">
```{r, fig.height = 7}
p <- germline_patient_data %>% 
  filter(!is.na(Race1)) %>%
  ggplot(aes(x=Race1, y=Age_at_MMonly_diagnosis), fill=Race1) + 
  geom_boxplot(color= c("#A92E5EFF", "#E65D2FFF")) +
  theme_minimal(base_size = 22) +
  labs(x=NULL, y="Age at Diagnosis", title="Age of MM diagnosis Distribution")
p + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means()
```
</div>
</div>

<br>

```{r tbl for number}
eth_age <- germline_patient_data %>% 
  mutate(Whole = "MM patients with WES") %>% 
  select(Ethnicity1, Age_at_MMonly_diagnosis) %>%
  tbl_summary(by = Ethnicity1, 
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% add_p()

race_age <- germline_patient_data %>% 
  mutate(Whole = "MM patients with WES") %>% 
  select(Race1, Age_at_MMonly_diagnosis) %>%
  tbl_summary(by = Race1, 
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% add_p()

raceeth_age <- germline_patient_data %>% 
  select(raceeth, Age_at_MMonly_diagnosis) %>%
  tbl_summary(by = raceeth, 
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% add_p()
```

Interestingly, we noticed a significant difference between disparities on the age at diagnosis. White Hispanics are diagnosed at a median age of `r inline_text(raceeth_age, variable = Age_at_MMonly_diagnosis, column = "White Non-Hispanic", pattern = "{median} years old")`. A significant earlier age at diagnosis is measured for Hispanics, `r inline_text(eth_age, variable = Age_at_MMonly_diagnosis, column = "Hispanic", pattern = "{median} years old")` , and for Blacks, `r inline_text(race_age, variable = Age_at_MMonly_diagnosis, column = "Black", pattern = "{median} years old")`, compare to Whites Hispanics (`r inline_text(race_age, variable = Age_at_MMonly_diagnosis, column = "p.value")`).


# III. RaceEthnicity

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
raceeth_surv <- germline_patient_data

mysurv <- Surv(time = raceeth_surv$month_at_os, event = raceeth_surv$os_event)
myplot <- survfit(mysurv~raceeth, data = raceeth_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = raceeth_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           palette = c("blue", "darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3))

tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Gender, ISS_grp, HCT_ever) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_os, 
                             event = raceeth_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_os, 
             event = raceeth_surv$os_event) ~ raceeth + Age_at_MMonly_diagnosis + Gender + ISS_grp + HCT_ever, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# eth_cox_os <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
# eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = raceeth_surv$month_at_progression_drug, event = raceeth_surv$drug_progression_event)
myplot <- survfit(mysurv~raceeth, data = raceeth_surv)
ggsurvplot(myplot, data = raceeth_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           palette = c("blue", "darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3))

tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Gender, ISS_grp, HCT_ever) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_progression_drug, 
                             event = raceeth_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_progression_drug, 
             event = raceeth_surv$drug_progression_event) ~ raceeth + Age_at_MMonly_diagnosis + Gender + ISS_grp + HCT_ever, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# eth_cox_pfs <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
# eth_cox_pfs
```
</div>
</div>

<br>

## CH & RaceEthnicity
### PFS
```{r, fig.height = 7}
mysurv <- Surv(time = raceeth_surv$month_at_progression_drug, event = raceeth_surv$drug_progression_event)
myplot <- survfit(mysurv~raceeth + CH_status, data = raceeth_surv)
ggsurvplot(myplot, data = raceeth_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "raceeth",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 3), linetype = guide_legend(nrow = 2))
```

```{r}
tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Gender, ISS_grp, HCT_ever, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_progression_drug, 
                             event = raceeth_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_progression_drug, 
             event = raceeth_surv$drug_progression_event) ~ raceeth + Age_at_MMonly_diagnosis + Gender + ISS_grp + HCT_ever + CH_status, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

```{r, fig.height = 7}
# mysurv <- Surv(time = raceeth_surv$month_at_progression_drug, event = raceeth_surv$drug_progression_event)
# myplot <- survfit(mysurv~raceeth + CH_status, data = raceeth_surv)
# ggsurvplot(myplot, data = raceeth_surv, facet.by = "MM_age_cat5",
#            title = "PFS",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(20, "bold", "black"),
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
# color = "raceeth",
#            linetype = "CH_status",
#            xlab = "Time in months"
# )+ guides(colour = guide_legend(nrow = 4))
```

```{r, fig.height = 7}
# mysurv <- Surv(time = raceeth_surv$month_at_progression_drug, event = raceeth_surv$drug_progression_event)
# myplot <- survfit(mysurv~raceeth + CH_status, data = raceeth_surv)
# ggsurvplot(myplot, data = raceeth_surv, facet.by = "MM_age_cat3",
#            title = "PFS",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(20, "bold", "black"),
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months"
# )+ guides(colour = guide_legend(nrow = 4))
```
<br>
<br>

### OS
```{r, fig.height = 7}
mysurv <- Surv(time = raceeth_surv$month_at_os, event = raceeth_surv$os_event)
myplot <- survfit(mysurv~raceeth + CH_status, data = raceeth_surv)
ggsurvplot(myplot, data = raceeth_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Race/Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "raceeth",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 3), linetype = guide_legend(nrow = 2))
```

```{r}
tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Gender, ISS_grp, HCT_ever, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_os, 
                             event = raceeth_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_os, 
                   event = raceeth_surv$os_event) ~ raceeth + Age_at_MMonly_diagnosis + Gender + ISS_grp + HCT_ever + CH_status, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```


***

# III. Ethnicity

## Summary

No difference in gender, ISS, CH and delay between diagnosis and treatment.
Most patients received drugs. Should we check the one who didn't to justify if they are just lost of follow up?
```{r}
germline_patient_data %>% 
  select(Age_at_MMonly_diagnosis,
         Gender, raceeth, Race, Ethnicity1, ISS, ISS_grp, ISS_at_MMdx, ISSdx_grp, CH_status,
         time_to_hct, time_to_drug,
         Drugs_ever, HCT_ever) %>%
  tbl_summary(by = Ethnicity1, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric", ISS_at_MMdx ~ "alphanumeric", ISSdx_grp ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05) %>% as_gt() %>%  
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```

## Survivals

No difference in OS within Ethnicity. The hazard ratio is decreased by 26% for Hispanic population but is not significant. Notably, age at diagnosis, ISS (ISS_grp) and receiving HCT are factor involved in OS in univariable and multivariable analysis. (Give data)
We observed that Hispanic population is doing better for PFS with a HR of progression increased by 53% compare to NH and 77% when adjusting by age at diagnosis, gender, ISS (ISS_grp) and receiving HCT
We noted that in multivariable analysis being a male increase significantly the HR of progression by 29% (add data, pval = 0.029). 
Age at diagnosis is not a factor involved in PFS but is in OS.
Not surprisingly ISS stage III have a worse survival (OS and PFS) and receiving HCT on the contrary increase it.


<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
germline_patient_surv <- germline_patient_data
ethnicity_surv <- germline_patient_surv %>% filter(!is.na(Ethnicity1))

mysurv <- Surv(time = ethnicity_surv$month_at_os, event = ethnicity_surv$os_event)
myplot <- survfit(mysurv~Ethnicity1, data = ethnicity_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = ethnicity_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           palette = c("darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
)

tbl1 <- ethnicity_surv %>% select(Ethnicity1, Age_at_MMonly_diagnosis, Gender, ISS, Drugs_ever, HCT_ever) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = ethnicity_surv$month_at_os, 
                             event = ethnicity_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = ethnicity_surv$month_at_os, 
             event = ethnicity_surv$os_event) ~ Ethnicity1 + Age_at_MMonly_diagnosis + Gender + ISS + Drugs_ever + HCT_ever, data =  ethnicity_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_os <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = ethnicity_surv$month_at_progression_drug, event = ethnicity_surv$drug_progression_event)
myplot <- survfit(mysurv~Ethnicity1, data = ethnicity_surv)
ggsurvplot(myplot, data = ethnicity_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           palette = c("darkred", "darkgreen"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

tbl1 <- ethnicity_surv %>% select(Ethnicity1, Age_at_MMonly_diagnosis, Gender, ISS, HCT_ever) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = ethnicity_surv$month_at_progression_drug, 
                             event = ethnicity_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = ethnicity_surv$month_at_progression_drug, 
             event = ethnicity_surv$drug_progression_event) ~ Ethnicity1 + Age_at_MMonly_diagnosis + Gender + ISS + HCT_ever, data =  ethnicity_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_pfs <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_pfs
```
</div>
</div>

<br>

Age is significantly associated with OS `r inline_text(tbl1, variable = Age_at_MMonly_diagnosis, pattern = "(OR {estimate}; 95% CI {conf.low}, {conf.high}; {p.value})")`.

<br>

I thought it could be a year effect compare to Taiga said (maybe earlier it wasn't a difference like he said but now that Hispanic may have more access to care we can see the difference) but no.

```{r}
mysurv <- Surv(time = ethnicity_surv$month_at_progression_drug, event = ethnicity_surv$drug_progression_event)
myplot <- survfit(mysurv~Ethnicity1+diagnosis_MM_year, data = ethnicity_surv)
ggsurvplot_facet(myplot, data = ethnicity_surv, facet.by = "diagnosis_MM_year",
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months",
           palette = c("darkred", "darkgreen"),
           conf.int = FALSE
)
```


## CH & Ethnicity in PFS

CH in not a factor in PFS as we observed a slight decrease of the HR of progression for patients with no CH
```{r, fig.height = 7}
mysurv <- Surv(time = ethnicity_surv$month_at_progression_drug, event = ethnicity_surv$drug_progression_event)
myplot <- survfit(mysurv~Ethnicity1 + CH_status, data = ethnicity_surv)
ggsurvplot(myplot, data = ethnicity_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "Ethnicity1",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 2))
```

```{r}
tbl1 <- ethnicity_surv %>% select(Ethnicity1, Age_at_MMonly_diagnosis, Gender, ISS, HCT_ever, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = ethnicity_surv$month_at_progression_drug, 
                             event = ethnicity_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = ethnicity_surv$month_at_progression_drug, 
             event = ethnicity_surv$drug_progression_event) ~ Ethnicity1 + Age_at_MMonly_diagnosis + Gender + ISS + HCT_ever + CH_status, data =  ethnicity_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

```{r, fig.height = 7}
# mysurv <- Surv(time = ethnicity_surv$month_at_progression_drug, event = ethnicity_surv$drug_progression_event)
# myplot <- survfit(mysurv~Ethnicity1 + CH_status, data = ethnicity_surv)
# ggsurvplot(myplot, data = ethnicity_surv, facet.by = "MM_age_cat5",
#            title = "PFS",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(20, "bold", "black"),
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months"
# )+ guides(colour = guide_legend(nrow = 4))
```

```{r, fig.height = 7}
# mysurv <- Surv(time = ethnicity_surv$month_at_progression_drug, event = ethnicity_surv$drug_progression_event)
# myplot <- survfit(mysurv~Ethnicity1 + CH_status, data = ethnicity_surv)
# ggsurvplot(myplot, data = ethnicity_surv, facet.by = "MM_age_cat3",
#            title = "PFS",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(20, "bold", "black"),
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months"
# )+ guides(colour = guide_legend(nrow = 4))
```
<br>
<br>

***

# IV. Race

## Summary

We observed a difference in gender but no difference in ISS, CH and if or type of treatment received.
Interestingly a longer delay to receive HCT was observed for Black population.

```{r demo in germline}
germline_patient_data %>% 
  select(Age_at_MMonly_diagnosis,
         Gender, raceeth, Race1, Ethnicity1, ISS, ISS_grp, ISS_at_MMdx, ISSdx_grp, CH_status,
         time_to_hct, time_to_drug,
         Drugs_ever, HCT_ever) %>%
  tbl_summary(by = Race1, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric", ISS_grp ~ "alphanumeric", ISS_at_MMdx ~ "alphanumeric", ISSdx_grp ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05) %>% as_gt() %>%  
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```


## Survivals

We observed that while PFS is not different between White and Black, Black shows a greater OS with a HR of death decreased by 49% (data) but this effect is not significant when adjusted by age, gender, ISS, and receiving a treatment. 
Age at diagnosis ISS and receiving HCT are factor associated with OS in the multivariable analysis.
In the multivariable analysis, Gender is associated to better progression free survival.

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
race_surv <- germline_patient_surv #%>% 
  # mutate(Race1 = factor(Race, levels=c("White", "Black")))  %>% filter(!is.na(Race))

mysurv <- Surv(time = race_surv$month_at_os, event = race_surv$os_event)
myplot <- survfit(mysurv~Race1, data = race_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Race1.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = race_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Race",
           # # legend.labs = c("White", "Black"),
           palette = c("blue", "lightsalmon1"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
)

tbl1 <- race_surv %>% select(Race1, Age_at_MMonly_diagnosis, Gender, ISS, Drugs_ever, HCT_ever) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = race_surv$month_at_os, 
                             event = race_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = race_surv$month_at_os, 
             event = race_surv$os_event) ~ Race1 + Age_at_MMonly_diagnosis + Gender + ISS + Drugs_ever + HCT_ever, data =  race_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
race_cox_os <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
race_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = race_surv$month_at_progression_drug, event = race_surv$drug_progression_event)
myplot <- survfit(mysurv~Race1, data = race_surv)
ggsurvplot(myplot, data = race_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Race",
           # # legend.labs = c("White", "Black"),
           palette = c("blue", "lightsalmon1"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

tbl1 <- race_surv %>% select(Race1, Age_at_MMonly_diagnosis, Gender, ISS, HCT_ever) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = race_surv$month_at_progression_drug, 
                             event = race_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = race_surv$month_at_progression_drug, 
             event = race_surv$drug_progression_event) ~ Race1 + Age_at_MMonly_diagnosis + Gender + ISS + HCT_ever, data =  race_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
race_cox_pfs <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
race_cox_pfs
```
</div>
</div>


















```{r}
mysurv <- Surv(time = race_surv$month_at_os, event = race_surv$os_event)
myplot <- survfit(mysurv~Race1+diagnosis_MM_year, data = race_surv)
ggsurvplot_facet(myplot, data = race_surv, facet.by = "diagnosis_MM_year",
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months",
           # legend = "top",
           # legend.title = "Ethnicity",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           # palette = c("darkred", "darkgreen"),
           # # color = "diagnosis_MM_year",
           # # linetype = "Ethnicity",
           # pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 3,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(5, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
)
```


## Black & CH

Very interestingly, we observed that while CH doesn't impact OS in white population, in Black population, presence of CH decrease drastically the survival rate by 80ish% (data).
Age at diagnosis, ISS are associated with OS in blacks alone and in combined White/Black population.

https://www.biorxiv.org/content/10.1101/782748v1.full.pdf
```{r, fig.height = 7}
mysurv <- Surv(time = race_surv$month_at_os, event = race_surv$os_event)
myplot <- survfit(mysurv~Race1+ CH_status, data = race_surv)

ggsurvplot(myplot, data = race_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "Race",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "Race1",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 2))

tbl1 <- race_surv %>% select(Race1, Age_at_MMonly_diagnosis, Gender, ISS, Drugs_ever, HCT_ever, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = race_surv$month_at_os, 
                             event = race_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = race_surv$month_at_os, 
             event = race_surv$os_event) ~ Race1 + Age_at_MMonly_diagnosis + Gender + ISS + Drugs_ever + HCT_ever + CH_status, data =  race_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
racech_cox_os <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
racech_cox_os
```


### Summary
```{r demo in black}
germline_patient_data %>% 
  select(Age_at_MMonly_diagnosis, diagnosis_MM_year, MM_age_cat3, MM_age_cat5,
         time_to_drug, Drugs_ever, HCT_ever, Gender, Race1, Ethnicity1, ISS, CH_status) %>%
  tbl_summary(by = CH_status, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2)) %>% bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05) %>% as_gt() %>%  
  gt::tab_source_note(gt::md("*ISS is reported for active MM patients only*"))
```



## In black

```{r}
# In black population only
race_black <- race_surv %>% filter(Race == "Black")
mysurv <- Surv(time = race_black$month_at_os, event = race_black$os_event)
myplot <- survfit(mysurv~Race1 + CH_status, data = race_black)
# jpeg(paste0(path, "/Figures/Moffitt Symposium/OS in Black CH.jpeg"), height = 600, width = 600)
ggsurvplot(myplot, data = race_black,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           color = "Race1",
           linetype = "CH_status",
           palette = c("lightsalmon1", "blue"),
           xlim = c(-15, 400),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (count(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           tables.theme = theme_survminer(font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
) + guides(linetype = guide_legend(nrow = 2, title = "")) + guides(colour = guide_legend(nrow = 2))
# dev.off()

tbl1 <- race_black %>% select(Age_at_MMonly_diagnosis, Gender, ISS, Drugs_ever, HCT_ever, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = race_black$month_at_os, 
                             event = race_black$os_event)),
                   exponentiate = TRUE,
                   label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis", CH_status ~ "CH status",
                                Drugs_ever ~ "Drugs ever", HCT_ever ~ "HCT ever")
                   ) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = race_black$month_at_os, 
                   event = race_black$os_event) ~ Age_at_MMonly_diagnosis + Gender + ISS + HCT_ever + CH_status, data =  race_black) %>%
  tbl_regression(exponentiate = TRUE,
                 label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis", CH_status ~ "CH status",
                              HCT_ever ~ "HCT ever")
                 ) %>% bold_p(t = .05)
tbl <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**")) %>% as_gt() %>% 
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
tbl
```
<br>

```{r, fig.height = 7}
# mysurv <- Surv(time = race_black$month_at_os, event = race_black$os_event)
# myplot <- survfit(mysurv~CH_status, data = race_black)
# ggsurvplot(myplot, data = race_black, facet.by = "MM_age_cat5",
#            title = "OS",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(20, "bold", "black"),
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months", 
#            legend.title = "",
#            palette = "race"
# )
```

```{r, fig.height = 7}
# mysurv <- Surv(time = race_black$month_at_os, event = race_black$os_event)
# myplot <- survfit(mysurv~CH_status, data = race_black)
# ggsurvplot(myplot, data = race_black, facet.by = "MM_age_cat3",
#            title = "O",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(20, "bold", "black"),
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months", 
#            legend.title = "",
#            palette = ""
# )

germline_patient_data %>% 
  select(diagnosis_MM_year, MM_age_cat3, MM_age_cat5, Race1, CH_status) %>%
  tbl_summary(by = Race1, 
              sort = list(everything() ~ "frequency")) %>%
  bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05) %>% as_gt() %>%  
  gt::tab_source_note(gt::md("*ISS is reported for active MM patients only*"))
```





<!-- #### -->
<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r, fig.height = 7} -->

<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r, fig.height = 7} -->

<!-- ``` -->
<!-- </div> -->
<!-- </div> -->
 

<br>
<br>
<!-- Leave a space here -->

# Others

***
## 4.CH

### Summary

We see **more male with CH** (66%) vs Female with CH (34%), more Female without CH (45%) vs Male without CH (55%) p=0.079.
CH is more represented in **severe ISS** (II-III).
<div class = "row">
<div class = "col-md-6">
```{r}

germline_patient_data %>%
  select(CH_status, ISS) %>% 
  tbl_summary(by = ISS, 
              sort = list(everything() ~ "frequency")) %>% add_p() %>% bold_p(t = .05) %>% bold_labels() %>% as_gt()
germline_patient_data %>%
  select(CH_status, ISS_grp) %>% 
  tbl_summary(by = ISS_grp, 
              sort = list(everything() ~ "frequency")) %>% add_p() %>% bold_p(t = .05) %>% bold_labels() %>% as_gt()
race_surv %>%
  select(CH_status, Race1) %>% 
  tbl_summary(by = Race1, 
              sort = list(everything() ~ "frequency")) %>% add_p() %>% bold_p(t = .05) %>% bold_labels() %>% as_gt()
germline_patient_data %>%
    filter(str_detect(Ethnicity, "Hispanic")) %>% 
  select(CH_status, Ethnicity) %>% 
  tbl_summary(by = Ethnicity, 
              sort = list(everything() ~ "frequency")) %>% add_p() %>% bold_p(t = .05) %>% bold_labels() %>% as_gt()
```

</div>

<div class = "col-md-6">
```{r}

```
</div>
</div>

### Survivals

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = germline_patient_surv)
# jpeg(paste0(path, "/Figures/Survivals/CHIP/OS by CH.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "", 
           # # legend.labs = c("CH", "No CH"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.17,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black"))
)
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = germline_patient_surv$month_at_progression_drug, event = germline_patient_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "", 
           # # legend.labs = c("CH", "No CH"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.17,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black"))
)
```
</div>
</div>

***
