---
title: "HCT Report"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cerulean #darkly
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 25px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")

```


<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">HCT</span>

</div>
<br>

```{r library, include=FALSE}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(viridis)
library(lubridate)
library(gtsummary)
library(survival)
library(survminer)
```

```{r load}
# Global_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Global_data.rds")
germline_patient_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/germline_patient_data.rds")
# Treatment1 <-  readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Treatment1.rds")
path <- fs::path("","Volumes","Gillis_Research","Christelle Colin-Leitzinger", "CHIP in Avatar")
# patients_removed_nonMM <- c("A000428", "A000456")
```


***
# I.Demographics of MM patients with WES who received HCT
## 1.General

```{r demographics in germline}
germline_patient_data <- germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  filter(!is.na(date_of_MMonly_diagnosis)) %>% 
  mutate(CH_status = str_replace(CH_status, "No_CH", "No CH"))

hct_data <- germline_patient_data %>% 
  filter(!is.na(date_of_bmt_1))

hct_data %>% 
  mutate(Whole = "HCT patients") %>% 
  mutate(Race = str_replace(Race, "Unknown", NA_character_)) %>% 
  mutate(Ethnicity = str_replace(Ethnicity, "Unknown", NA_character_)) %>% 
  select(Age_at_MMonly_diagnosis, 
         Gender, Race, Ethnicity, Whole, ISS) %>%
  tbl_summary(by = Whole, label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis"),
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels() %>% as_gt() %>%  
  gt::tab_source_note(gt::md("**Data for active MM patients who received HCT**")) %>% 
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```

## 2.Disease status at time of bood collection (`r emo::ji("warning")` is not at time of HCT)
```{r}
hct_data %>%
  mutate(Whole = "HCT patients") %>% 
  select(Disease_Status_germline, Whole) %>% 
  tbl_summary(by = Whole, label = list(Disease_Status_germline ~ "Disease Status at Time of Bood Collection")) %>% 
  bold_labels() %>% as_gt()  %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```

## 3.Time from blood collection to hct
```{r}
hct_data %>%
  mutate(days_at_hct_blood = (interval(start= hct_data$collectiondt_germline, end= hct_data$date_of_bmt_1)  /  duration(n=1, unit="days"))) %>% 
  mutate(Whole = "HCT patients") %>% 
  select(days_at_hct_blood, Whole, time_to_hct) %>% 
  tbl_summary(by = Whole, label = list(days_at_hct_blood ~ "Time to HCT from time to Bood Collection (in days)", time_to_hct ~ "Time to HCT from MM diagnosis (in days)")) %>% 
  bold_labels() %>% as_gt() %>%  
  gt::tab_source_note(gt::md("**HCT happen 694 days before blood collection**")) %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

hct_data %>% 
  mutate(HCT_vs_germline = str_replace(HCT_vs_germline, "No HCT", NA_character_)) %>% 
  select(HCT_vs_germline, time_to_hct) %>% 
  tbl_summary(label = list(time_to_hct ~ "Time to HCT (days)"),
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% as_gt() %>%  
  # gt::tab_source_note(gt::md("*ISS is reported for MM patients only*")) %>%  
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

hct_data %>% 
  mutate(days_at_hct_blood = (interval(start= hct_data$collectiondt_germline, end= hct_data$date_of_bmt_1)  /  duration(n=1, unit="days"))) %>% 
  mutate(HCT_vs_germline = str_replace(HCT_vs_germline, "No HCT", NA_character_)) %>% 
  select(HCT_vs_germline, days_at_hct_blood, HCT_vs_germline) %>% 
  tbl_summary(by = HCT_vs_germline, label = list(days_at_hct_blood ~ "Time to HCT from time to Bood Collection (in days)"),
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% as_gt() %>%  
  # gt::tab_source_note(gt::md("*ISS is reported for MM patients only*")) %>%  
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

```

## 4.Year at HCT
```{r}
hct_data %>% mutate(year_at_first_bmt = as.factor(year(date_of_bmt_1))) %>% 
  mutate(Whole = "HCT patients") %>% 
  select(year_at_first_bmt, Whole) %>% 
  tbl_summary(by = Whole) %>% 
  bold_labels() %>% as_gt()  %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```

## 5.Name of first regimen for HCT patients
```{r}
hct_data %>% 
  mutate(Whole = "HCT patients") %>% 
  select(first_regimen_name, Whole) %>% 
  tbl_summary(by = Whole,
              sort = list(everything() ~ "frequency")) %>% 
  bold_labels() %>% as_gt()  %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

hct_data %>%
  select(avatar_id, starts_with("treatment_line_")) %>% 
  pivot_longer(cols = starts_with("treatment_line_"),
               names_to = "regimen_number", values_to = "treatment_line_", values_drop_na = TRUE) %>% 
  filter(treatment_line_ < 90) %>% 
  arrange(desc(treatment_line_)) %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(treatment_line_ = case_when(
    treatment_line_ > 5         ~ "more than 5",
    TRUE                        ~ as.character(treatment_line_)
  )) %>% 
  select(treatment_line_) %>% 
  tbl_summary(label = list(treatment_line_ ~ "Number of line received by HCT patients"),
              missing_text = "No Drugs") %>% bold_labels() %>% as_gt() %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

hct_data %>%
  select(avatar_id, date_of_bmt_1, starts_with("line_start")) %>% 
  pivot_longer(cols = starts_with("line_start"),
               names_to = "line", values_to = "date", values_drop_na = TRUE) %>% 
  filter(date < date_of_bmt_1) %>% 
  group_by(avatar_id) %>% 
  mutate(line_bf_hct = n()) %>% 
  ungroup() %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  select(line_bf_hct) %>% 
  tbl_summary(label = list(line_bf_hct ~ "Number of line of drugs received by patients before HCT"),
              missing_text = "No Drugs") %>% bold_labels() %>% as_gt() %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```

<br>

<!-- #### ISS -->
<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r, fig.height = 7} -->
<!-- germline_patient_surv <- germline_patient_data %>% distinct(avatar_id, .keep_all = TRUE) -->
<!-- mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event) -->
<!-- myplot <- survfit(mysurv~ISS, data = germline_patient_surv) -->
<!-- # jpeg(paste0(path, "/Figures/Survivals/Demographic/OS ISS.jpeg"), width = 1200, height = 900) -->
<!-- ggsurvplot(myplot, data = germline_patient_surv, -->
<!--            title = "OS ISS from date of diagnosis", -->
<!--            font.main = c(24, "bold", "black"), -->
<!--            font.x = c(20, "bold", "black"), -->
<!--            font.y = c(20, "bold", "black"), -->
<!--            font.legend = c(20, "bold", "black"), -->
<!--            font.tickslab = c(18, "bold", "black"), -->
<!--            size = 1.5, -->

<!--            xlab = "Time in months",  -->
<!--            legend = "top", -->
<!--            legend.title = "ISS", -->
<!--            # # legend.labs = c("No ISS", "ISS"), -->
<!--            # palette = c("darkred", "darkgreen", "grey"), -->
<!--            pval = TRUE, -->
<!--            conf.int = FALSE, -->
<!--            # Add risk table -->
<!--            tables.height = 0.3, -->
<!--            risk.table.title = "Risk table (number(%))", -->
<!--            risk.table = "abs_pct", -->
<!--            risk.table.y.text = FALSE, -->
<!--            risk.table.fontsize = 4, -->
<!--            tables.theme = theme_survminer(base_size = 5, -->
<!--                                           font.main = c(16, "bold", "black"), -->
<!--                                           font.x = c(16, "bold", "black"), -->
<!--                                           font.y = c(16, "bold", "transparent"), -->
<!--                                           font.tickslab = c(19, "bold", "black")), -->
<!--            # Censor -->
<!--            censor = TRUE -->
<!-- ) -->
<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r, fig.height = 7} -->
<!-- mysurv <- Surv(time = germline_patient_surv$month_at_progression_Dx, event = germline_patient_surv$Progression_event) -->
<!-- myplot <- survfit(mysurv~ISS, data = germline_patient_surv) -->
<!-- # jpeg(paste0(path, "/Figures/Survivals/Demographic/PFS ISS from Dx.jpeg"), width = 1200, height = 900) -->
<!-- ggsurvplot(myplot, data = germline_patient_surv, -->
<!--            title = "PFS ISS from Dx", -->
<!--            font.main = c(24, "bold", "black"), -->
<!--            font.x = c(20, "bold", "black"), -->
<!--            font.y = c(20, "bold", "black"), -->
<!--            font.legend = c(20, "bold", "black"), -->
<!--            font.tickslab = c(18, "bold", "black"), -->
<!--            size = 1.5, -->

<!--            xlab = "Time in months",  -->
<!--            legend = "top", -->
<!--            legend.title = "ISS", -->
<!--            # # legend.labs = c("No ISS", "ISS"), -->
<!--            # palette = c("darkred", "darkgreen", "grey"), -->
<!--            pval = TRUE, -->
<!--            conf.int = FALSE, -->
<!--            # Add risk table -->
<!--            tables.height = 0.3, -->
<!--            risk.table.title = "Risk table (number(%))", -->
<!--            risk.table = "abs_pct", -->
<!--            risk.table.y.text = FALSE, -->
<!--            risk.table.fontsize = 4, -->
<!--            tables.theme = theme_survminer(base_size = 5, -->
<!--                                           font.main = c(16, "bold", "black"), -->
<!--                                           font.x = c(16, "bold", "black"), -->
<!--                                           font.y = c(16, "bold", "transparent"), -->
<!--                                           font.tickslab = c(19, "bold", "black") -->
<!--            ), -->
<!--            # Censor -->
<!--            censor = TRUE -->
<!-- ) -->
<!-- # dev.off() -->

<!-- mysurv <- Surv(time = germline_patient_surv$month_at_progression_drug, event = germline_patient_surv$drug_progression_event) -->
<!-- myplot <- survfit(mysurv~ISS, data = germline_patient_surv) -->
<!-- # jpeg(paste0(path, "/Figures/Survivals/Demographic/PFS ISS from drugs date.jpeg"), width = 1200, height = 900) -->
<!-- ggsurvplot(myplot, data = germline_patient_surv, -->
<!--            title = "PFS ISS from drug date", -->
<!--            font.main = c(24, "bold", "black"), -->
<!--            font.x = c(20, "bold", "black"), -->
<!--            font.y = c(20, "bold", "black"), -->
<!--            font.legend = c(20, "bold", "black"), -->
<!--            font.tickslab = c(18, "bold", "black"), -->
<!--            size = 1.5, -->

<!--            xlab = "Time in months",  -->
<!--            legend = "top", -->
<!--            legend.title = "ISS", -->
<!--            # # legend.labs = c("No ISS", "ISS"), -->
<!--            # palette = c("darkred", "darkgreen", "grey"), -->
<!--            pval = TRUE, -->
<!--            conf.int = FALSE, -->
<!--            # Add risk table -->
<!--            tables.height = 0.3, -->
<!--            risk.table.title = "Risk table (number(%))", -->
<!--            risk.table = "abs_pct", -->
<!--            risk.table.y.text = FALSE, -->
<!--            risk.table.fontsize = 4, -->
<!--            tables.theme = theme_survminer(base_size = 5, -->
<!--                                           font.main = c(16, "bold", "black"), -->
<!--                                           font.x = c(16, "bold", "black"), -->
<!--                                           font.y = c(16, "bold", "transparent"), -->
<!--                                           font.tickslab = c(19, "bold", "black") -->
<!--            ), -->
<!--            # Censor -->
<!--            censor = TRUE -->
<!-- ) -->
<!-- ``` -->
<!-- </div> -->
<!-- </div> -->



***
## 6.CH
CH status is assessed by using the Avatar WES data.
```{r}
hct_data %>%
  distinct(avatar_id, .keep_all = TRUE) %>% 
  select(CH_status) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>% 
  bold_labels() %>% as_gt() %>%  
  gt::tab_style(
    style = gt::cell_text(
      color = "#0099CC"
    ),
    locations = gt::cells_column_labels(everything())
  ) 
```
<br>
<br>

***
# II.Survivals by HCT
All survival analysis are limited to the germline population.

## 1.HCT `r emo::ji("warning")` here it includes patients with no HCT
pfs-hct is when HCT happened before germline = Yes whereas HCT_ever is a more general Yes/No HCT.
<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
germline_patient_surv <- germline_patient_data

mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)
myplot <- survfit(mysurv~HCT_ever, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS HCT",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
           xlim = c(-15, 400),
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 2))



mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)
myplot <- survfit(mysurv~HCT_vs_germline, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS HCT_vs_germline",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3))

mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)
myplot <- survfit(mysurv~HCT_before_germline, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS HCT_before_germline",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 2))

```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
germline_patient_surv %>% select(HCT_ever) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = germline_patient_surv$month_at_os, 
                             event = germline_patient_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()

germline_patient_surv %>% select(HCT_vs_germline) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = germline_patient_surv$month_at_os, 
                             event = germline_patient_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()


```
</div>
</div>

<span style="color:red">
**It is a difference of survival between patient who received HCT BEFORE GERMLINE or not but it may be due to the stage...**
</span>


<br>
<br>

## 2.HCT & ISS

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 9}
germline_patient_surv <- germline_patient_data %>% mutate(HCT_vs_germline = str_replace(HCT_vs_germline, "No HCT", NA_character_))

mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)

myplot <- survfit(mysurv~HCT_ever+ISS, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS HCT",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("No Treatment", "Treatment"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "ISS",
           linetype = "HCT_ever",
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
) + guides(linetype = guide_legend(nrow = 2, title = "")) + guides(colour = guide_legend(nrow = 3))

myplot <- survfit(mysurv~HCT_vs_germline+ISS, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS HCT",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("No Treatment", "Treatment"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "ISS",
           linetype = "HCT_vs_germline",
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
) + guides(linetype = guide_legend(nrow = 3, title = "")) + guides(colour = guide_legend(nrow = 3))

```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
tbl1 <- germline_patient_surv %>% select(HCT_vs_germline, ISS) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = germline_patient_surv$month_at_os, 
                             event = germline_patient_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = germline_patient_surv$month_at_os, 
             event = germline_patient_surv$os_event) ~ HCT_vs_germline + ISS, data =  germline_patient_surv) %>%
  tbl_regression(exponentiate = TRUE)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))

tbl1 <- germline_patient_surv %>% select(HCT_ever, ISS) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = germline_patient_surv$month_at_os, 
                             event = germline_patient_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = germline_patient_surv$month_at_os, 
             event = germline_patient_surv$os_event) ~ HCT_ever + ISS, data =  germline_patient_surv) %>%
  tbl_regression(exponentiate = TRUE)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
</div>
</div>

<span style="color:red">
**When compare survival between patient who received SCT by ISS stage, it is no difference**
</span>

<br>
<br>

## 3.HCT & CH
<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
germline_patient_surv <- germline_patient_data %>% distinct(avatar_id, .keep_all = TRUE)
mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)
myplot <- survfit(mysurv~HCT_ever+CH_status, data = germline_patient_surv)
# jpeg(paste0(path, "/Figures/Survivals/CHIP/OS HCT by CH.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS HCT ever",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"), #20
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "HCT_ever",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 2))

mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)
myplot <- survfit(mysurv~HCT_vs_germline+CH_status, data = germline_patient_surv)
# jpeg(paste0(path, "/Figures/Survivals/CHIP/OS HCT by CH.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS HCT_vs_germline",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"), #20
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "HCT_vs_germline",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 2))

myplot <- survfit(mysurv~HCT_before_germline+CH_status, data = germline_patient_surv)
# jpeg(paste0(path, "/Figures/Survivals/CHIP/OS HCT by CH.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS HCT_before_germline",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"), #20
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "HCT_before_germline",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 2))

```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
tbl1 <- germline_patient_surv %>% select(HCT_before_germline, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = germline_patient_surv$month_at_os, 
                             event = germline_patient_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = germline_patient_surv$month_at_os, 
             event = germline_patient_surv$os_event) ~ HCT_before_germline + CH_status, data =  germline_patient_surv) %>%
  tbl_regression(exponentiate = TRUE)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))

tbl1 <- germline_patient_surv %>% select(HCT_ever, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = germline_patient_surv$month_at_os, 
                             event = germline_patient_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = germline_patient_surv$month_at_os, 
             event = germline_patient_surv$os_event) ~ HCT_ever + CH_status, data =  germline_patient_surv) %>%
  tbl_regression(exponentiate = TRUE)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
</div>
</div>

<span style="color:red">
**Interestingly this difference of survival between +/- SCT is seen in patients with no CH.**
</span>

<br>
<br>


## 4.HCT +/- drugs `r emo::ji("warning")` here it includes patients with no HCT
<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)
myplot <- survfit(mysurv~pfs_treatment, data = germline_patient_surv)
# jpeg(paste0(path, "/Figures/Survivals/Treatment/OS by Treatment.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS Treatment",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # legend.labs = c("No Treatment", "Treatment"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 2))
# dev.off()

myplot <- survfit(mysurv~pfs_treatment+CH_status, data = germline_patient_surv)
# jpeg(paste0(path, "/Figures/Survivals/CHIP/OS Treatment CH.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS Treatment",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # legend.labs = c("No Treatment", "Treatment"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "pfs_treatment",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(linetype = guide_legend(nrow = 2, title = "")) + guides(colour = guide_legend(nrow = 2))
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = germline_patient_surv$month_at_progression_drug, event = germline_patient_surv$drug_progression_event)
myplot <- survfit(mysurv~pfs_treatment, data = germline_patient_surv)
# jpeg(paste0(path, "/Figures/Survivals/Treatment/PFS by Treatment from treatment date.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "PFS Treatment from drug date",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("No Treatment", "Treatment"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 2))
# dev.off()

myplot <- survfit(mysurv~pfs_treatment+CH_status, data = germline_patient_surv)
# jpeg(paste0(path, "/Figures/Survivals/CHIP/PFS by Treatment CH from treatment date.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "PFS Treatment from drug date",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("No Treatment", "Treatment"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "pfs_treatment",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-15, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
) + guides(linetype = guide_legend(nrow = 2, title = "")) + guides(colour = guide_legend(nrow = 2))
```
</div>
</div>

<br>
<br>