---
title: "SM Report"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: no
    theme: cerulean
    highlight: pygments
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{=html}
<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 25px;
    margin-bottom: 25px;
}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")

```

```{=html}
<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
```
::: {.blue}
[Smoldering]{style="color: white;"}
:::

<br>

```{r library, include=FALSE}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(viridis)
library(lubridate)
library(gtsummary)
library(survival)
library(survminer)
```

```{r load}
# Global_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Global_data.rds")
germline_patient_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/germline_patient_data.rds")
# Treatment1 <-  readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Treatment1.rds")
# path <- fs::path("","Volumes","Gillis_Research","Christelle Colin-Leitzinger", "CHIP in Avatar")
# patients_removed_nonMM <- c("A000428", "A000456")
```


***

# Smoldering patients and CH
<br>
<br>

```{r}
# germline_patient_data %>% 
#   mutate(Whole = "SM patients") %>% 
# 
#   select(smoldering_status, Whole) %>%
#   tbl_summary(by = Whole, 
#               sort = list(everything() ~ "frequency"),
#               missing = "no") %>% 
#   bold_labels() %>% as_gt() %>%  
#   gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```

***

## 1. Demographics

```{r data}
sm_patient_data <- germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(smoldering_status = factor(smoldering_status, levels = c("Is Smoldering", "Progressed from Smoldering"))) %>% 
  filter(!is.na(smoldering_status))
rm(germline_patient_data)
```

```{r}
sm_patient_data %>% 
  select(Age_at_MMonly_diagnosis, 
         Gender, Race, Ethnicity, smoldering_status, ISS, ISS_grp) %>%
  tbl_summary(by = smoldering_status, label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis"),
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric", ISS_grp ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels() %>% add_p(correct = FALSE) %>% bold_p(t = .05) %>% as_gt() %>%  
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```

As you can see Smoldering diagnosis is messed up because we see patients that are coded as still smolderimg in the data but Raghu gave a disease status at Blood collection with a MM status.
```{r}
sm_patient_data %>%
  select(Disease_Status_germline, smoldering_status) %>% 
  tbl_summary(by = smoldering_status, 
              label = list(Disease_Status_germline ~ "Disease Status at Time of Blood Collection"),
              sort = list(everything() ~ "frequency")) %>% 
  bold_labels() %>% as_gt()  %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```
<br>
<br>

***

## 3. CH 
```{r}
sm_patient_data %>%
  select(smoldering_status, CH_status) %>%
  tbl_summary(by = smoldering_status,
              sort = list(everything() ~ "frequency"),
              missing = "no") %>%
  bold_labels() %>% add_p(correct = FALSE) %>% bold_p(t = .05) %>% as_gt() %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```
```{r}
# sm_patient_data %>% 
#   select(smoldering_status, CH_status) %>%
#   tbl_summary(by = smoldering_status, 
#               sort = list(everything() ~ "frequency"),
#               missing = "no") %>% 
#   bold_labels() %>% add_p(correct = FALSE) %>% bold_p(t = .05) %>% as_hux_table(strip_md_bold = TRUE)#%>% as_gt() %>%  
#   # gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```
```{r}
# sm_patient_data %>% 
#   select(smoldering_status, CH_status) %>%
#   tbl_summary(by = smoldering_status, 
#               sort = list(everything() ~ "frequency"),
#               missing = "no") %>% 
#   bold_labels() %>% add_p(correct = FALSE) %>% bold_p(t = .05) %>% as_flextable(return_calls = FALSE)
```
```{r}
# sm_patient_data %>% 
#   select(smoldering_status, CH_status) %>%
#   tbl_summary(by = smoldering_status, 
#               sort = list(everything() ~ "frequency"),
#               missing = "no") %>% 
#   bold_labels() %>% add_p(correct = FALSE) %>% bold_p(t = .05) %>% as_flex_table(strip_md_bold = TRUE) %>% flextable::border()
```
<br>
<br>

***

## 4. Drugs

```{r}
sm_patient_data %>% 
  select(first_regimen_name, smoldering_status) %>% 
  tbl_summary(by = smoldering_status,
              label = list(first_regimen_name ~ "Drugs receive as first regimen at all time"),
              sort = list(everything() ~ "frequency")) %>% 
  bold_labels() %>% as_gt()  %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```

```{r}
sm_patient_data %>% 
  mutate(ab = case_when(
    line_start_date_1 >= date_of_MM_diagnosis    ~ "remove",
    line_start_date_1 < date_of_MM_diagnosis     ~ "keep",
    TRUE ~ NA_character_
  )) %>% 
  filter(ab == "keep") %>% 
  select(first_regimen_name, smoldering_status, imids_maintenance) %>% 
  tbl_summary(by = smoldering_status,
              label = list(first_regimen_name ~ "Drugs receive as first regimen BEFORE MM diagnosis"),
              sort = list(everything() ~ "frequency")) %>% 
  bold_labels() %>% as_gt()  %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```
<br>
<br>

***

## Few KM

I only did OS because I will probably need to recode progression date differently as my progression variable is currently looking at progression AFTER MM diagnosis.  

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
# mysurv <- Surv(time = sm_patient_data$month_at_os, event = sm_patient_data$os_event)
# myplot <- survfit(mysurv~CH_status, data = sm_patient_data)
# # jpeg(paste0(path, "/Figures/Survivals/CHIP/OS Drugs by CH.jpeg"), width = 1200, height = 900)
# ggsurvplot(myplot, data = sm_patient_data,
#            title = "OS Drugs ever",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(10, "bold", "black"), #20
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months", 
#            legend = "top",
#            legend.title = "",
#            # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
#            # palette = c("darkred", "darkgreen", "grey"),
#            pval = TRUE,
#            conf.int = FALSE,
# # Add risk table
#            tables.height = 0.3,
#            risk.table.title = "Risk table (number(%))",
#            risk.table = "abs_pct",
#            risk.table.y.text = FALSE,
#            risk.table.fontsize = 4,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")),
#            # Censor
#            censor = TRUE
# ) + guides(colour = guide_legend(ncol = 1))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
# tbl1 <- sm_patient_data %>% select(CH_status) %>% 
#   tbl_uvregression(method = survival::coxph, 
#                    y = (Surv(time = sm_patient_data$month_at_os, 
#                              event = sm_patient_data$os_event)),
#                    exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
#   bold_labels() %>% italicize_levels()
# tbl2 <- coxph(Surv(time = sm_patient_data$month_at_os, 
#              event = sm_patient_data$os_event) ~ CH_status, data =  sm_patient_data) %>%
#   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
:::
:::

<br>

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
# mysurv <- Surv(time = sm_patient_data$month_at_progression_drug, event = sm_patient_data$drug_progression_event)
# myplot <- survfit(mysurv~CH_status, data = sm_patient_data)
# ggsurvplot(myplot, data = sm_patient_data,
#            title = "PFS Drugs from  DRUGS date",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(10, "bold", "black"), # 20
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months", 
#            legend = "top",
#            legend.title = "",
#            # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
#            # palette = c("darkred", "darkgreen", "grey"),
#            pval = TRUE,
#            conf.int = FALSE,
#            # Add risk table
#            tables.height = 0.3,
#            risk.table.title = "Risk table (number(%))",
#            risk.table = "abs_pct",
#            risk.table.y.text = FALSE,
#            risk.table.fontsize = 4,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            # Censor
#            # censor = TRUE
# ) + guides(colour = guide_legend(ncol = 1))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
# tbl1 <- sm_patient_data %>% select(CH_status) %>%
#   tbl_uvregression(method = survival::coxph,
#                    y = (Surv(time = sm_patient_data$month_at_progression_drug,
#                              event = sm_patient_data$drug_progression_event)),
#                    exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>%
#   bold_labels() %>% italicize_levels()
# tbl2 <- coxph(Surv(time = sm_patient_data$month_at_progression_drug,
#              event = sm_patient_data$drug_progression_event) ~ CH_status, data =  sm_patient_data) %>%
#   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
:::
:::

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
mysurv <- Surv(time = sm_patient_data$month_at_os, event = sm_patient_data$os_event)
myplot <- survfit(mysurv~CH_status+smoldering_status, data = sm_patient_data)
# jpeg(paste0(path, "/Figures/Survivals/CHIP/OS Drugs by CH.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = sm_patient_data,
           title = "OS Drugs ever",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"), #20
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
           # xlim = c(-10, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
tbl1 <- sm_patient_data %>% select(CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = sm_patient_data$month_at_os, 
                             event = sm_patient_data$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = sm_patient_data$month_at_os, 
             event = sm_patient_data$os_event) ~ CH_status, data =  sm_patient_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
:::
:::

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
# sm_surv <- sm_patient_data %>% filter(str_detect(smoldering_status, "Pro"))
# mysurv <- Surv(time = sm_surv$month_at_os, event = sm_surv$os_event)
# myplot <- survfit(mysurv~CH_status+smoldering_status, data = sm_surv)
# # jpeg(paste0(path, "/Figures/Survivals/CHIP/OS Drugs by CH.jpeg"), width = 1200, height = 900)
# ggsurvplot(myplot, data = sm_surv,
#            title = "OS Drugs ever",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(10, "bold", "black"), #20
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months", 
#            legend = "top",
#            legend.title = "",
#            # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
#            # palette = c("darkred", "darkgreen", "grey"),
#            pval = TRUE,
#            conf.int = FALSE,
#            # xlim = c(-10, 400),
# # Add risk table
#            tables.height = 0.3,
#            risk.table.title = "Risk table (number(%))",
#            risk.table = "abs_pct",
#            risk.table.y.text = FALSE,
#            risk.table.fontsize = 4,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")),
#            # Censor
#            censor = TRUE
# ) + guides(colour = guide_legend(ncol = 1))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
# tbl1 <- sm_surv %>% select(CH_status) %>% 
#   tbl_uvregression(method = survival::coxph, 
#                    y = (Surv(time = sm_surv$month_at_os, 
#                              event = sm_surv$os_event)),
#                    exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
#   bold_labels() %>% italicize_levels()
# tbl2 <- coxph(Surv(time = sm_surv$month_at_os, 
#              event = sm_surv$os_event) ~ CH_status, data =  sm_surv) %>%
#   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
:::
:::

<br>

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
# mysurv <- Surv(time = sm_patient_data$month_at_progression_drug, event = sm_patient_data$drug_progression_event)
# myplot <- survfit(mysurv~CH_status+smoldering_status, data = sm_patient_data)
# ggsurvplot(myplot, data = sm_patient_data,
#            title = "PFS Drugs from  DRUGS date",
#            font.main = c(24, "bold", "black"),
#            font.x = c(20, "bold", "black"),
#            font.y = c(20, "bold", "black"),
#            font.legend = c(10, "bold", "black"), # 20
#            font.tickslab = c(18, "bold", "black"),
#            size = 1.5,
#            
#            xlab = "Time in months", 
#            legend = "top",
#            legend.title = "",
#            # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
#            # palette = c("darkred", "darkgreen", "grey"),
#            pval = TRUE,
#            conf.int = FALSE,
#            # Add risk table
#            tables.height = 0.3,
#            risk.table.title = "Risk table (number(%))",
#            risk.table = "abs_pct",
#            risk.table.y.text = FALSE,
#            risk.table.fontsize = 4,
#            tables.theme = theme_survminer(base_size = 5,
#                                           font.main = c(16, "bold", "black"),
#                                           font.x = c(16, "bold", "black"),
#                                           font.y = c(16, "bold", "transparent"),
#                                           font.tickslab = c(19, "bold", "black")
#            ),
#            # Censor
#            # censor = TRUE
# ) + guides(colour = guide_legend(ncol = 1))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
# tbl1 <- sm_patient_data %>% select(CH_status) %>%
#   tbl_uvregression(method = survival::coxph,
#                    y = (Surv(time = sm_patient_data$month_at_progression_drug,
#                              event = sm_patient_data$drug_progression_event)),
#                    exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>%
#   bold_labels() %>% italicize_levels()
# tbl2 <- coxph(Surv(time = sm_patient_data$month_at_progression_drug,
#              event = sm_patient_data$drug_progression_event) ~ CH_status, data =  sm_patient_data) %>%
#   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
:::
:::

<br> 

***