---
title: "Drugs Report"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cerulean #darkly
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{=html}
<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 25px;
    margin-bottom: 25px;
}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")

```

```{=html}
<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
```
::: {.blue}
[Drugs]{style="color: white;"}
:::

<br>

```{r library, include=FALSE}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(viridis)
library(lubridate)
library(gtsummary)
library(survival)
library(survminer)
```

```{r load}
# Global_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Global_data.rds")
germline_patient_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/germline_patient_data.rds")
# Treatment1 <-  readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Treatment1.rds")
# path <- fs::path("","Volumes","Gillis_Research","Christelle Colin-Leitzinger", "CHIP in Avatar")
# patients_removed_nonMM <- c("A000428", "A000456")
```


***

# MM patients with WES who received Drugs
<br> 
<br>

***

## 1. Demographics

```{r demographics in germline}
germline_patient_data <- germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  filter(!is.na(date_of_MMonly_diagnosis)) %>% 
  mutate(CH_status = str_replace(CH_status, "No_CH", "No CH"))

drugs_data <- germline_patient_data %>% 
  filter(!is.na(line_start_date_1)) %>% 
  mutate(days_at_Drugs_blood = interval(start= collectiondt_germline, end= date_of_bmt_1)  /  duration(n=1, unit="days")) %>% 
    mutate(time_to_Drugs = interval(start= collectiondt_germline, end= date_of_MMonly_diagnosis)  /  duration(n=1, unit="days"))


drugs_data %>% 
  mutate(Whole = "Drugs patients") %>% 
  mutate(Race = str_replace(Race, "Unknown", NA_character_)) %>% 
  mutate(Ethnicity = str_replace(Ethnicity, "Unknown", NA_character_)) %>% 
  select(Age_at_MMonly_diagnosis, 
         Gender, Race, Ethnicity, Whole, ISS, ISS_grp) %>%
  tbl_summary(by = Whole, label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis"),
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric", ISS_grp ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels() %>% as_gt() %>%  
  gt::tab_source_note(gt::md("**Data for active MM patients who received Drugs**")) %>% 
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```
<br> 
<br>

***

## 2. Disease Status at Time of Blood Collection (`r emo::ji("warning")` is not at time of Drugs)

```{r}
drugs_data %>%
  mutate(Whole = "Drugs patients") %>% 
  select(Disease_Status_germline, Whole) %>% 
  tbl_summary(by = Whole, 
              label = list(Disease_Status_germline ~ "Disease Status at Time of Blood Collection"),
              sort = list(everything() ~ "frequency")) %>% 
  bold_labels() %>% as_gt()  %>%
  gt::tab_source_note(gt::md("**All these patients are now active MM patients**")) %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```
<br> 
<br>

***

## 3. Time from blood collection to Drugs

```{r, fig.width=10}
drugs_data %>%
  mutate(Whole = "Drugs patients") %>% 
  select(days_at_Drugs_blood, Whole, time_to_Drugs) %>% 
  tbl_summary(by = Whole, label = list(days_at_Drugs_blood ~ "Time to Drugs from Blood Collection (in days)", time_to_Drugs ~ "Time to Drugs from MM diagnosis (in days)")) %>% 
  bold_labels() %>% as_gt() %>%  
  gt::tab_source_note(gt::md("**Drugs happen 694 days before blood collection**")) %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

p <- qplot(x =days_at_Drugs_blood, data=subset(drugs_data,!is.na(days_at_Drugs_blood)), fill=..count.., 
           geom="histogram", 
           binwidth = 100,
           ) 
p + scale_fill_viridis_c(
  alpha = 1,
  begin = 0,
  end = 1,
  direction = 1,
  option = "D",
  values = NULL,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
) +
  theme_minimal(base_size = 14) +
  labs(x="Time to Drugs from Blood Collection (in days)", y="Number of Patient", title="Number of Days Between Blood Collection and Drugs",
       caption = "Each bar represents 100 days")+
  
  annotate("text", x = -3750, y = 38, # x and y coordinates of the text
           label = "Drugs before \nblood collection", 
           size = 4) +
  annotate("segment", x = -7500, xend = -15, y = 35, yend = 35,
           size = 1, colour = "steelblue")+
  annotate("text", x = 1450, y = 38, # x and y coordinates of the text
           label = "Drugs after \nblood collection", 
           size = 4) +
  annotate("segment", x = 15, xend = 2500, y = 35, yend = 35,
           size = 1, colour = "blue") + 
  
  scale_x_continuous(breaks=c(-7500, -6225, -5000, -3775, -2500, -1225, 0, 200, 1225, 2500),
          minor_breaks=c(0, 200)
          ) + 
  theme(axis.text.x = element_text(angle=90, size=c(12, 12, 12, 12, 12, 12, 9, 12, 12)))# +
  # geom_text(stat="count", aes(label=stat(count)), vjust=-0.2) 

drugs_data %>% 
  # mutate(drugs_before_germline = str_replace(drugs_before_germline, "No Drugs", NA_character_)) %>% 
  select(drugs_before_germline) %>% 
  tbl_summary(label = list(drugs_before_germline ~ "Drugs vs germline"),
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% as_gt() %>%  
  # gt::tab_source_note(gt::md("*ISS is reported for MM patients only*")) %>%  
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

drugs_data %>% 
  # mutate(days_at_Drugs_blood = (interval(start= drugs_data$collectiondt_germline, end= drugs_data$date_of_bmt_1)  /  duration(n=1, unit="days"))) %>% 
  mutate(drugs_before_germline = str_replace(drugs_before_germline, "No Drugs", NA_character_)) %>% 
  select(drugs_before_germline, days_at_Drugs_blood, drugs_before_germline) %>% 
  tbl_summary(by = drugs_before_germline, label = list(days_at_Drugs_blood ~ "Time to Drugs from Blood Collection (in days)"),
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% as_gt() %>%  
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

drugs_data %>% 
  # mutate(days_at_Drugs_blood = (interval(start= drugs_data$collectiondt_germline, end= drugs_data$date_of_bmt_1)  /  duration(n=1, unit="days"))) %>% 
  # filter(days_at_Drugs_blood < 0) %>% 
  mutate(drugs_before_germline = str_replace(drugs_before_germline, "No Drugs", NA_character_)) %>%
  select(drugs_before_germline, Disease_Status_germline) %>% 
  tbl_summary(by = drugs_before_germline, label = list(Disease_Status_germline ~ "Disease Status at Blood collection"),
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% as_gt() %>%  
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

```

<br> 
<br>

***

## 4. Name of first regimen for Drugs patients

The first table include the name of the **FIRST regimen** name.  

```{r}
drugs_data %>% 
  mutate(Whole = "Drugs patients") %>% 
  select(first_regimen_name, Whole) %>% 
  tbl_summary(by = Whole,
              sort = list(everything() ~ "frequency")) %>% 
  bold_labels() %>% as_gt()  %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

drugs_data %>%
  select(avatar_id, starts_with("treatment_line_")) %>% 
  pivot_longer(cols = starts_with("treatment_line_"),
               names_to = "regimen_number", values_to = "treatment_line_", values_drop_na = TRUE) %>% 
  filter(treatment_line_ < 90) %>% 
  arrange(desc(treatment_line_)) %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(treatment_line_ = case_when(
    treatment_line_ > 5         ~ "more than 5",
    TRUE                        ~ as.character(treatment_line_)
  )) %>% 
  select(treatment_line_) %>% 
  tbl_summary(label = list(treatment_line_ ~ "Number of line received"),
              missing_text = "No Drugs") %>% bold_labels() %>% as_gt() %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

drugs_data %>%
  select(avatar_id, collectiondt_germline, starts_with("line_start")) %>% 
  pivot_longer(cols = starts_with("line_start"),
               names_to = "line", values_to = "date", values_drop_na = TRUE) %>% 
  filter(date < collectiondt_germline) %>% 
  group_by(avatar_id) %>% 
  mutate(line_bf_Drugs = n()) %>% 
  ungroup() %>% 
  arrange(desc(line_bf_Drugs)) %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  select(line_bf_Drugs) %>% 
  tbl_summary(label = list(line_bf_Drugs ~ "Number of line of drugs received by patients before Blood"),
              missing_text = "No Drugs") %>% bold_labels() %>% as_gt() %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```
<br> 
<br>

***

## 6. CH

CH status is assessed by using the Avatar WES data.

```{r}
drugs_data %>%
  distinct(avatar_id, .keep_all = TRUE) %>% 
  select(CH_status) %>% 
  tbl_summary(sort = list(everything() ~ "frequency")) %>% 
  bold_labels() %>% as_gt() %>%  
  gt::tab_style(
    style = gt::cell_text(
      color = "#0099CC"
    ),
    locations = gt::cells_column_labels(everything())
  ) 

tbl1 <- drugs_data %>% 
  mutate(Race = str_replace(Race, "Unknown", NA_character_)) %>% 
  mutate(Ethnicity = str_replace(Ethnicity, "Unknown", NA_character_)) %>% 
  select(Age_at_MMonly_diagnosis, 
         Gender, Race, Ethnicity, ISS, ISS_grp, CH_status) %>%
  tbl_summary(by = CH_status, label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis"),
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)# %>% as_gt() %>%  
  # gt::tab_source_note(gt::md("**Data for active MM patients who received Drugs**")) %>% 
  # gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

tbl2 <- drugs_data %>%
  select(avatar_id, collectiondt_germline, starts_with("line_start"), CH_status) %>% 
  pivot_longer(cols = starts_with("line_start"),
               names_to = "line", values_to = "date", values_drop_na = TRUE) %>% 
  filter(date < collectiondt_germline) %>% 
  group_by(avatar_id) %>% 
  mutate(line_bf_blood = n()) %>% 
  ungroup() %>% 
  distinct(avatar_id, .keep_all = TRUE) %>%
  mutate(line_bf_blood = ifelse(line_bf_blood > 5, "greater than 5", as.character(line_bf_blood))) %>% 
  mutate(line_bf_blood = factor(line_bf_blood)) %>% 
  select(line_bf_blood, CH_status) %>% 
  tbl_summary(by = CH_status,
              label = list(line_bf_blood ~ "Number of line of drugs received by patients before blood collection"),
              missing_text = "No Drugs") %>% 
  
  # modify_spanning_header(everything() ~ "Number of line of drugs received by patients before blood collection") %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05) #%>% as_gt() %>% 
  # gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))

tbl_stack(list(tbl1, tbl2)) %>% as_gt() %>%
  gt::tab_source_note(gt::md("**Data for active MM patients who received Drugs**")) %>%
  gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
```

```{r}
# drugs_data %>%
#   select(avatar_id, starts_with("treatment_line_"), CH_status) %>% 
#   pivot_longer(cols = starts_with("treatment_line_"),
#                names_to = "regimen_number", values_to = "treatment_line_", values_drop_na = TRUE) %>% 
#   filter(treatment_line_ < 90) %>% 
#   arrange(desc(treatment_line_)) %>% 
#   distinct(avatar_id, .keep_all = TRUE) %>% 
#   mutate(treatment_line_ = case_when(
#     as.numeric(treatment_line_) > 5         ~ "more than 5",
#     TRUE                                    ~ as.character(treatment_line_)
#   )) %>% 
#   select(treatment_line_, CH_status) %>% 
#   tbl_summary(by = treatment_line_,
#               missing_text = "No Drugs") %>%
#   modify_spanning_header(everything() ~ "Number of line received by Drugs patients") %>% 
#   bold_labels() %>% add_p() %>% bold_p(t = .05) %>% as_gt() %>% 
#   gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))
# 
# drugs_data %>%
#   select(avatar_id, date_of_bmt_1, starts_with("line_start"), CH_status) %>% 
#   pivot_longer(cols = starts_with("line_start"),
#                names_to = "line", values_to = "date", values_drop_na = TRUE) %>% 
#   filter(date < date_of_bmt_1) %>% 
#   group_by(avatar_id) %>% 
#   mutate(line_bf_Drugs = n()) %>% 
#   ungroup() %>% 
#   distinct(avatar_id, .keep_all = TRUE) %>% 
#   select(line_bf_Drugs, CH_status) %>% 
#   tbl_summary(by = line_bf_Drugs,
#               missing_text = "No Drugs") %>% 
#   modify_spanning_header(everything() ~ "Number of line of drugs received by patients before Drugs") %>% 
#   bold_labels() %>% add_p() %>% bold_p(t = .05) %>% as_gt() %>% 
#   gt::tab_style(style = gt::cell_text(color = "#0099CC"), locations = gt::cells_column_labels(everything()))


```

```{r}
# drugs_data %>% 
#   distinct(avatar_id, .keep_all = TRUE) %>% 
#   # mutate(Whole = "Germline patients") %>% 
#   select(ISS, ISS_grp CH_status) %>%
#   tbl_summary(by = CH_status, 
#               sort = list(everything() ~ "frequency", ISS ~ "alphanumeric")) %>%
#   bold_labels() %>% add_p() %>% bold_p(t = .05) %>% as_gt() %>% 
#   gt::tab_source_note(gt::md("**Data on Drugs patients**"))

```
<br> 
<br>

***

## 7. Survivals by Drugs

All survival analysis are limited to the germline population.

### Drugs `r emo::ji("warning")` here it includes patients with no Drugs

`r emo::ji("pad")` **For OS the time point "zero" is diagnosis.**  

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
germline_patient_surv <- germline_patient_data

mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)
myplot <- survfit(mysurv~Drugs_ever, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS Drugs",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
           xlim = c(-10, 400),
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 2))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
germline_patient_surv %>% select(Drugs_ever) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = germline_patient_surv$month_at_os, 
                             event = germline_patient_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
```
:::
:::

<br>

Also `r emo::ji("question_mark")` to do PFS, we need the time point "zero" for the "No Drugs" population.  

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
mysurv <- Surv(time = germline_patient_surv$month_at_progression_drug, event = germline_patient_surv$drug_progression_event)
myplot <- survfit(mysurv~Drugs_ever, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "PFS Drugs from  DRUGS date",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"), # 20
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           # censor = TRUE
)
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
# germline_patient_surv %>% select(Drugs_ever) %>% 
#   tbl_uvregression(method = survival::coxph, 
#                    y = (Surv(time = germline_patient_surv$month_at_progression_drug, 
#                              event = germline_patient_surv$drug_progression_event)),
#                    exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
#   bold_labels() %>% italicize_levels()
```
:::
:::

<br> 
<br>

### Drugs & ISS

::: {.row}
::: {.col-md-6}
```{r, fig.height = 9}
germline_patient_surv <- germline_patient_data %>% mutate(drugs_before_germline = str_replace(drugs_before_germline, "No Drugs", NA_character_))

mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)

myplot <- survfit(mysurv~Drugs_ever+ISS, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS Drugs",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("No Treatment", "Treatment"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "ISS",
           linetype = "Drugs_ever",
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-10, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
) + guides(linetype = guide_legend(nrow = 2, title = "")) + guides(colour = guide_legend(nrow = 3))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
tbl1 <- germline_patient_surv %>% select(Drugs_ever, ISS) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = germline_patient_surv$month_at_os, 
                             event = germline_patient_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = germline_patient_surv$month_at_os, 
             event = germline_patient_surv$os_event) ~ Drugs_ever + ISS, data =  germline_patient_surv) %>%
  tbl_regression(exponentiate = TRUE)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
:::
:::

[ **Receiving Drugs increases survival in each ISS group**
]{style="color:red"}

<br> <br>

### Drugs & CH

#### In WES Population

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
germline_patient_surv <- germline_patient_data %>% distinct(avatar_id, .keep_all = TRUE)
mysurv <- Surv(time = germline_patient_surv$month_at_os, event = germline_patient_surv$os_event)
myplot <- survfit(mysurv~Drugs_ever+CH_status, data = germline_patient_surv)
# jpeg(paste0(path, "/Figures/Survivals/CHIP/OS Drugs by CH.jpeg"), width = 1200, height = 900)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "OS Drugs ever",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"), #20
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "Drugs_ever",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-10, 400),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 2))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
tbl1 <- germline_patient_surv %>% select(Drugs_ever, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = germline_patient_surv$month_at_os, 
                             event = germline_patient_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = germline_patient_surv$month_at_os, 
             event = germline_patient_surv$os_event) ~ Drugs_ever + CH_status, data =  germline_patient_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
:::
:::

<br>

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
mysurv <- Surv(time = germline_patient_surv$month_at_progression_drug, event = germline_patient_surv$drug_progression_event)
myplot <- survfit(mysurv~Drugs_ever+CH_status, data = germline_patient_surv)
ggsurvplot(myplot, data = germline_patient_surv,
           title = "PFS Drugs from  DRUGS date",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"), # 20
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           color = "Drugs_ever",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           # censor = TRUE
) + guides(colour = guide_legend(nrow = 2))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
# tbl1 <- germline_patient_surv %>% select(Drugs_ever, CH_status) %>% 
#   tbl_uvregression(method = survival::coxph, 
#                    y = (Surv(time = germline_patient_surv$month_at_progression_drug, 
#                              event = germline_patient_surv$drug_progression_event)),
#                    exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
#   bold_labels() %>% italicize_levels()
# tbl2 <- coxph(Surv(time = germline_patient_surv$month_at_progression_drug, 
#              event = germline_patient_surv$drug_progression_event) ~ Drugs_ever + CH_status, data =  germline_patient_surv) %>%
#   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
:::
:::

<br> 
<br>

#### In WES patient who had Blood collected before drugs

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
blood_bf_Drugs <- drugs_data %>% 
  filter(drugs_before_germline == "Blood before Drugs")

mysurv <- Surv(time = blood_bf_Drugs$month_at_os, event = blood_bf_Drugs$os_event)
myplot <- survfit(mysurv~CH_status, data = blood_bf_Drugs)
ggsurvplot(myplot, data = blood_bf_Drugs,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
                      xlim = c(-10, 150),
# Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")),
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3), linetype = guide_legend(nrow = 3))



```
:::

::: {.col-md-6}
```{r, fig.height = 7}
blood_bf_Drugs %>% select(CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = blood_bf_Drugs$month_at_os, 
                             event = blood_bf_Drugs$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
```
:::
:::

::: {.row}
::: {.col-md-6}
```{r, fig.height = 7}
mysurv <- Surv(time = blood_bf_Drugs$month_at_progression_drug, event = blood_bf_Drugs$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = blood_bf_Drugs)
ggsurvplot(myplot, data = blood_bf_Drugs,
           title = "PFS Drugs from  DRUGS date",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(10, "bold", "black"), # 20
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hipanic", "Non-Hispanic", "Unknown"),
           # palette = c("darkred", "darkgreen", "grey"),
           pval = TRUE,
           conf.int = FALSE,
           xlim = c(-10, 150),
           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           # censor = TRUE
) + guides(colour = guide_legend(nrow = 2))
```
:::

::: {.col-md-6}
```{r, fig.height = 7}
tbl1 <- blood_bf_Drugs %>% select(CH_status) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = blood_bf_Drugs$month_at_progression_drug,
                             event = blood_bf_Drugs$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = blood_bf_Drugs$month_at_progression_drug,
             event = blood_bf_Drugs$drug_progression_event) ~ CH_status, data =  blood_bf_Drugs) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariate**", "**Multivariate**"))
```
:::
:::


