---
title: "analysis added up after Taiga's comments"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: cerulean #darkly
    highlight: pygments
    # keep_md: true
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   <!-- margin-bottom: 100px; -->
}

table {
    margin-top: 25px;
    <!-- margin-bottom: 100px !important; -->
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")

```


```{r library, include=FALSE}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(viridis)
library(lubridate)
library(gtsummary)
library(survival)
library(survminer)
```

```{r load}
Global_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Global_data.rds")
germline_patient_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/germline_patient_data.rds")
Treatment1 <-  readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Treatment1.rds")
path <- fs::path("","Volumes","Gillis_Research","Christelle Colin-Leitzinger", "CHIP in Avatar")
patients_removed_nonMM <- c("A000428", "A000456")
```
<br>
<br>
<!-- Leave a space here -->

***
# I. Time to treatment or HCT start
From MM diagnosis so will be for patients who were MM or became MM (list of patients who became MM in IV)
```{r}
germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(time_to_hct, time_to_drug, Ethnicity) %>%
  tbl_summary(by = Ethnicity, 
              sort = list(everything() ~ "frequency"),
              digits = list(c(time_to_hct, time_to_drug) ~ 2)) %>% 
  bold_labels() %>% add_p()

germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(time_to_hct, time_to_drug, ISS) %>%
  tbl_summary(by = ISS, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(c(time_to_hct, time_to_drug) ~ 2)) %>% 
  bold_labels() %>% add_p()
```


# II. The patient with HCT without drug
A025764

# III. Drugs

## 1. Summary lines per patients
The max is 15 regimen
```{r}
germline_patient_data %>% 
  mutate(regimen_count = case_when(
    !is.na(line_start_date_1) &
      is.na(line_start_date_2)        ~ "1",
    !is.na(line_start_date_2) &
      is.na(line_start_date_3)        ~ "2",
    !is.na(line_start_date_3) &
      is.na(line_start_date_4)        ~ "3",
    !is.na(line_start_date_4) &
      is.na(line_start_date_5)        ~ "4",
    !is.na(line_start_date_5) &
      is.na(line_start_date_6)        ~ "5",
    !is.na(line_start_date_6) &
      is.na(line_start_date_7)        ~ "6",
    !is.na(line_start_date_7) &
      is.na(line_start_date_8)        ~ "7",
    !is.na(line_start_date_8) &
      is.na(line_start_date_9)        ~ "8",
    !is.na(line_start_date_9) &
      is.na(line_start_date_10)        ~ "9",
    !is.na(line_start_date_10) &
      is.na(line_start_date_11)        ~ "10",
    !is.na(line_start_date_11) &
      is.na(line_start_date_12)        ~ "11",
    !is.na(line_start_date_12) &
      is.na(line_start_date_13)        ~ "12",
    !is.na(line_start_date_13) &
      is.na(line_start_date_14)        ~ "13",
    !is.na(line_start_date_14) &
      is.na(line_start_date_15)        ~ "14",
    !is.na(line_start_date_15)         ~ "15"
  )) %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(regimen_count, Whole) %>%
  tbl_summary(by = Whole, 
              sort = list(everything() ~ "frequency")) %>% 
  bold_labels() %>% add_p()
```



## 2.Tables drugs separated by regimen 1,2,3
Include all germline patients. You can see in IV the drugs that the patients who were Smol and became MM had.
<div class = "row">
<div class = "col-md-4">
```{r, fig.height = 7}
germline_patient_data %>%
  mutate(Whole = "All regimen in 1st regimen") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  select(first_regimen_name, Whole) %>% 
  tbl_summary(by = Whole,
              sort = list(everything() ~ "frequency"),
              missing_text = "No Drugs") %>% bold_labels() %>%  as_gt()
```
</div>

<div class = "col-md-4">
```{r, fig.height = 7}
germline_patient_data %>%
  mutate(Whole = "All regimen in 2nd regimen") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  select(regimen_name_2, Whole) %>% 
  tbl_summary(by = Whole,
              sort = list(everything() ~ "frequency"),
              missing_text = "No regimen_name_2") %>% bold_labels() %>%  as_gt()
```
</div>

<div class = "col-md-4">
```{r, fig.height = 7}
germline_patient_data %>%
  mutate(Whole = "All regimen in 3rd regimen") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  select(regimen_name_3, Whole) %>% 
  tbl_summary(by = Whole,
              sort = list(everything() ~ "frequency"),
              missing_text = "No regimen_name_3") %>% bold_labels() %>%  as_gt()
```
</div>
</div>


# IV. How many of the MGUS, Sm progressed to MM
```{r}
rmarkdown::paged_table(
germline_patient_data %>% 
  select(avatar_id, Disease_Status_germline, Dx_date_closest_germline, date_of_MMonly_diagnosis, first_regimen_name, date_of_bmt_1) %>% 
  filter(str_detect(Disease_Status_germline, "Mgus|Smoldering")) %>% 
  arrange(date_of_MMonly_diagnosis, Disease_Status_germline)
)
```

I definitely need to remove the non MM patients but I think I need to dig moe into these patients who had HCT but are not MM.

# V. Patients who got HCT by ISS
```{r}
germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(pfs_hct, HCT_ever, ISS) %>%
  tbl_summary(by = ISS, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric")) %>% 
  bold_labels() %>% add_p()

germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(pfs_hct, HCT_ever, ISS_grp) %>%
  tbl_summary(by = ISS_grp, 
              sort = list(everything() ~ "frequency", ISS_grp ~ "alphanumeric")) %>% 
  bold_labels() %>% add_p()

```

# VI. Things I look this morning before the meeting

## 1. Hispanic
```{r}
germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(Age_at_diagnosis_closest_germline, Gender, Race, Ethnicity, Whole, ISS) %>%
  tbl_summary(by = Ethnicity, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(c(Age_at_diagnosis_closest_germline, Race) ~ 2)) %>% 
  bold_labels() %>% add_p()

germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(Age_at_diagnosis_closest_germline, Ethnicity, Whole, ISS) %>%
  tbl_summary(by = ISS, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(c(Age_at_diagnosis_closest_germline) ~ 2)) %>% 
  bold_labels() %>% add_p()
```

## 2. Does CH status affect PFS in number?
```{r}
germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(Age_at_diagnosis_closest_germline, month_at_os, month_at_progression_drug, month_at_progression_hct, month_at_progression_rad, CH_status, ISS) %>%
  tbl_summary(by = CH_status, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(c(Age_at_diagnosis_closest_germline) ~ 2)) %>% 
  bold_labels() %>% add_p()

tbl1 <- germline_patient_data %>% 
  filter(CH_status == "CH") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(Age_at_diagnosis_closest_germline, month_at_os, month_at_progression_drug, month_at_progression_hct, month_at_progression_rad, CH_status, ISS) %>%
  tbl_summary(by = ISS, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(c(Age_at_diagnosis_closest_germline) ~ 2)) %>% 
  bold_labels() %>% add_p()

tbl2 <- germline_patient_data %>% 
  filter(CH_status == "No_CH") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(Age_at_diagnosis_closest_germline, month_at_os, month_at_progression_drug, month_at_progression_hct, month_at_progression_rad, CH_status, ISS) %>%
  tbl_summary(by = ISS, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(c(Age_at_diagnosis_closest_germline) ~ 2)) %>% 
  bold_labels() %>% add_p()

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**CH**", "**No_CH**"))

germline_patient_data %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(CH_status, ISS) %>%
  tbl_summary(by = ISS, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric")) %>% 
  bold_labels() %>% add_p()

tbl1 <- germline_patient_data %>% 
  filter(ISS == "I") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(Age_at_diagnosis_closest_germline, month_at_os, month_at_progression_drug, month_at_progression_hct, month_at_progression_rad, CH_status, ISS) %>%
  tbl_summary(by = CH_status, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(everything() ~ 2)) %>% 
  bold_labels() %>% add_p()

tbl2 <- germline_patient_data %>% 
  filter(ISS == "II") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(Age_at_diagnosis_closest_germline, month_at_os, month_at_progression_drug, month_at_progression_hct, month_at_progression_rad, CH_status, ISS) %>%
  tbl_summary(by = CH_status, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(everything() ~ 2)) %>% 
  bold_labels() %>% add_p()

tbl3 <- germline_patient_data %>% 
  filter(ISS == "III") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  mutate(Whole = "Germline patients") %>% 
  select(Age_at_diagnosis_closest_germline, month_at_os, month_at_progression_drug, month_at_progression_hct, month_at_progression_rad, CH_status, ISS) %>%
  tbl_summary(by = CH_status, 
              sort = list(everything() ~ "frequency", ISS ~ "alphanumeric"),
              digits = list(everything() ~ 2)) %>% 
  bold_labels() %>% add_p()

tbl_merge(list(tbl1, tbl2, tbl3), tab_spanner = c("**ISS I**", "**ISS II**", "**ISS III**"))
```

<!-- ### -->
<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r, fig.height = 7} -->

<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r, fig.height = 7} -->

<!-- ``` -->
<!-- </div> -->
<!-- </div> -->
