---
title: "Sensitivity analysis"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: no
    theme: cerulean
    highlight: pygments
editor_options:
  chunk_output_type: console
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   margin-bottom: 25px;
}

table {
    margin-top: 25px;
    <!-- margin-bottom: 100px !important; -->
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 12, data_row.padding = gt::px(1))")

```


<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">Sensitivity analysis ISS with "Unknown" category</span>

</div>
<br>

```{r library, include=FALSE}
library(tidyverse)
library(data.table)
library(VennDiagram)
library(viridis)
library(lubridate)
library(gtsummary)
library(survival)
library(survminer)
```

```{r load}
# Global_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/Global_data.rds")
germline_patient_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/germline_patient_data.rds")
path <- ("/Users/colinccm/Documents/GitHub/Gillis/no vpn data/CHIP in Avatar")

germline_patient_data <- germline_patient_data %>% 
  rename(Sex = Gender) %>% 
  filter(is_patient_MM == "Yes" & is_MMDx_close_to_blood == "Yes") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  filter(raceeth != "Others") %>% 
  mutate(raceeth = factor(raceeth, levels=c("White Non-Hispanic", "Hispanic", "Black"))) %>% 
  mutate(ISS_at_MMdx = replace_na(ISS_at_MMdx, "Unknown")) %>%
  mutate(regimen_categoryVCD_1 = factor(regimen_categoryVCD_1, levels = c(
    "PI + IMIDs", "PI", "IMIDs", "Others"
  )))
```
<br>
<br>

I removed the "Others" in raceeth.


***

# Introduction

The ORIEN Avatar^TM Research Program focuses on patients with advanced primary or metastatic disease, those with limited treatment options, as well as patients who are likely to develop progressive disease, and will capitalize on the disease-area expertise of ORIEN's research scientists.   
It generates massive amounts of genetic and clinical information on patients consenting to the Total Cancer Care® (TCC) Protocol.   
ORIEN Avatar tracks patient clinical progressions over their lifetime to provide the richest data source for targeted cancer therapy advancement.  ORIEN Avatar data solution is optimized for oncology research by providing integrated longitudinal clinical data and whole exome, transcriptome and germline sequencing around consented oncology patients.   
We are focusing on patients with multiple myeloma disease.
We will focus on patient with active MM who had WES analyzed on their blood and CH were analyzed.

<br>

***

# I. MM Avatar data description


```{r}
germline_patient_data <- germline_patient_data %>% 
  mutate(HCT_comparison = case_when(
    is.na(date_of_bmt_1)              ~ "No HCT",
    time_to_hct < 314                 ~ "Early HCT",
    time_to_hct >= 314                ~ "Late HCT"
  )) %>% 
  mutate(HCT_comparison = factor(HCT_comparison, levels = c("No HCT", "Early HCT", "Late HCT")))

contamination_tumor_ids <- paste("A025767", "A025771", "A007368", "A007322", "A000111", sep = "|")
```

<div class = "row">
<div class = "col-md-5">
```{r}
germ_data <- germline_patient_data %>% 
  mutate(Whole = "MM patients with WES") %>% 
  select(Age_at_MMonly_diagnosis, Sex, raceeth, Whole, ISS_at_MMdx) %>%
  tbl_summary(by = Whole, 
              label = list(raceeth ~ "Race/ethnicity", Age_at_MMonly_diagnosis ~ "Age at MM diagnosis"),
              sort = list(everything() ~ "frequency", ISS_at_MMdx ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2)#,
              # missing = "no"
              ) %>% 
  bold_labels()
germ_data
```
</div>

<div class = "col-md-7">
```{r}
p <- qplot(x =Age_at_MMonly_diagnosis, data=subset(germline_patient_data,!is.na(Age_at_MMonly_diagnosis)), fill=..count.., geom="histogram") 
p + scale_fill_viridis_c(
  alpha = 1,
  begin = 0,
  end = 1,
  direction = 1,
  option = "D",
  values = NULL,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
) +
  theme_minimal(base_size = 16) +
  labs(x="Age at Diagnosis", y="Number of Patient", title="Repartition of Age at Diagnosis in MM Avatar")
```
</div>
</div>

<br>

Patients represented in MM Avatar data are characterized by a median age at diagnosis of `r inline_text(germ_data, variable = Age_at_MMonly_diagnosis, column = "MM patients with WES", pattern = "{median} years old")`. We saw slightly more male `r inline_text(germ_data, variable = Sex, level = "Male", column = "MM patients with WES")`, a majority of White Non-Hispanic `r inline_text(germ_data, variable = raceeth, level = "White Non-Hispanic", column = "MM patients with WES")`(Blacks `r inline_text(germ_data, variable = raceeth, level = "Black", column = "MM patients with WES")`, Hispanic `r inline_text(germ_data, variable = raceeth, level = "Hispanic", column = "MM patients with WES")`).
A higher proportion of patients have a lower clinical stage at blood with `r inline_text(germ_data, variable = ISS_at_MMdx, level = "I", column = "MM patients with WES")` patients with ISS I.

<br>
<br>

***

# II. Age Distribution in Disparities

## Raceeth
```{r}
# tiff(paste0(path, "/Disparities manuscript/age distribution boxplot raceeth.tiff"))
p <- ggplot(germline_patient_data, 
            aes(x=raceeth, y=Age_at_MMonly_diagnosis)) + 
  geom_boxplot(color = c("#03051AFF", "blue", "red"),
               alpha = 0.3,
               fill = c("#03051AFF", "blue", "red")) + 
  theme_minimal(base_size = 23) +
  labs(x=NULL, y="Age at Diagnosis", title="Age of MM Diagnosis Distribution")
p + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means(label.x = 1.75)
# dev.off()
```

```{r tbl for number}
raceeth_age <- germline_patient_data %>% 
  select(raceeth, Age_at_MMonly_diagnosis) %>%
  tbl_summary(by = raceeth, 
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)
```
<br>

Interestingly, we noticed a significant difference between disparities on the age at diagnosis. White Hispanics are diagnosed at a median age of `r inline_text(raceeth_age, variable = Age_at_MMonly_diagnosis, column = "White Non-Hispanic", pattern = "{median} years old")`. A significant earlier age at diagnosis is measured for Hispanics, `r inline_text(raceeth_age, variable = Age_at_MMonly_diagnosis, column = "Hispanic", pattern = "{median} years old")` , and for Blacks, `r inline_text(raceeth_age, variable = Age_at_MMonly_diagnosis, column = "Black", pattern = "{median} years old")`, compare to Whites Hispanics (`r inline_text(raceeth_age, variable = Age_at_MMonly_diagnosis, column = "p.value")`).

<br>
<br>

***

# III. RaceEthnicity

## Clinical

The p val for the regimen because R is running out of space.
If I run it by hand with `fisher.test(tbl1$raceeth, tbl1$regimen_category_1)`, I have the same.  
I can increase the wrospace `fisher.test(tbl1$raceeth, tbl1$regimen_category_1, workspace=2e9)`which give a good result, we should use that.
But it is a problem in the `gt summary` package which doesn't compute it even if I increse the workspace.
I am going to ask the person who build the package but for now I will give you the p val on the side.

```{r}
tbl1 <-
  germline_patient_data %>% 
  select(Age_at_MMonly_diagnosis, Sex, raceeth, ISS_at_MMdx,
         CH_status,
         HCT_ever, regimen_categoryVCD_1) %>%
  tbl_summary(by = raceeth, 
              label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis",
                           ISS_at_MMdx ~ "ISS at MM diagnosis",
                           HCT_ever ~ "HCT ever",
                           CH_status ~ "CH status"),
              sort = list(everything() ~ "frequency", ISS_at_MMdx ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2)
              ) %>% 
  bold_labels() %>% add_p(test.args = regimen_categoryVCD_1 ~ list(workspace=2e9)) %>% bold_p(t = .05)
```


```{r}
tbl2 <- germline_patient_data %>% 
  select(raceeth,
         time_to_drug, time_to_hct) %>%
  tbl_summary(by = raceeth, 
              label = list(time_to_drug ~ "Time to drug (days)", time_to_hct ~ "Time to HCT (days)"),
              sort = list(everything() ~ "frequency"),
              missing = "no"
              ) %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)

tbl_stack(list(tbl1, tbl2))
```

```{r}
# germline_patient_data %>% 
#   select(time_to_drug, HCT_comparison, raceeth) %>%
#   tbl_summary(by = raceeth, 
#               label = list(time_to_drug ~ "Time to drug (days)", HCT_comparison ~ "Time to HCT (days)"),
#               sort = list(everything() ~ "frequency"),
#               missing = "no") %>% bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
```

<br>
<br>

## Survivals

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
raceeth_surv <- germline_patient_data

mysurv <- Surv(time = raceeth_surv$month_at_os, event = raceeth_surv$os_event)
myplot <- survfit(mysurv~raceeth, data = raceeth_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
# pdf(paste0(path, "/Disparities manuscript/OS raceeth.pdf"))
ggsurvplot(myplot, data = raceeth_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           legend.labs = c("White Non-Hispanic", "Hispanic", "Black"),
           palette = c("#03051AFF", "blue", "red"),
           pval = TRUE,
           conf.int = FALSE,
           # xlim = c(-15, 400),


           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3))
# dev.off()

tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_os, 
                             event = raceeth_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_os, 
             event = raceeth_surv$os_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# eth_cox_os <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
# eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = raceeth_surv$month_at_progression_drug, event = raceeth_surv$drug_progression_event)
myplot <- survfit(mysurv~raceeth, data = raceeth_surv)
# pdf(paste0(path, "/Disparities manuscript/PFS raceeth.pdf"))

ggsurvplot(myplot, data = raceeth_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           legend.labs = c("White Non-Hispanic", "Hispanic", "Black"),
           palette = c("#03051AFF", "blue", "red"),
           pval = TRUE,
           conf.int = FALSE,
           # xlim = c(-10, 250),


           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3))
# dev.off()

tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_progression_drug, 
                             event = raceeth_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_progression_drug, 
             event = raceeth_surv$drug_progression_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# eth_cox_pfs <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
# eth_cox_pfs
```
</div>
</div>

<br>
<br>

***

# IV. CH

From here, remove 5 patients with tumor contamination

```{r}
patients_clean_tumor <- germline_patient_data %>% 
  filter(!str_detect(avatar_id, contamination_tumor_ids))
```

```{r}
ch_char <- patients_clean_tumor %>% 
  mutate(Whole = "MM patients at time of blood with WES") %>% 
  select(Whole, CH_status) %>%
  tbl_summary(by = Whole, 
              label = list(CH_status ~ "CH status"),
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels()
ch_char
```
<br>
<br>

Whole Exome Sequencing data identified CH mutations in `r inline_text(ch_char, variable = CH_status, level = "CH", column = "MM patients at time of blood with WES")`.
+ Add bioinfo results (oncoplot?).  
Type of genes affected...

<br>
<br>

```{r}
# ch_samples <- patients_clean_tumor %>%
#   select(CH_status) %>% 
#   tbl_summary(sort = list(everything() ~ "frequency")) %>% 
#   bold_labels()
# ch_samples
patients_clean_tumor %>%
  select(Age_at_MMonly_diagnosis, 
         Sex, raceeth, ISS_at_MMdx, ISS_at_MMdx,
         CH_status) %>% 
  tbl_summary(by = CH_status, 
              label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis", CH_status ~ "CH status"),
              sort = list(everything() ~ "frequency", ISS_at_MMdx ~ "alphanumeric", ISS_at_MMdx ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              percent = "row",
              missing = "no") %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)

ch_demo <- patients_clean_tumor %>%
  select(Age_at_MMonly_diagnosis, 
         Sex, raceeth, ISS_at_MMdx, ISS_at_MMdx,
         CH_status) %>% 
  tbl_summary(by = CH_status, 
              label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis", CH_status ~ "CH status"),
              sort = list(everything() ~ "frequency", ISS_at_MMdx ~ "alphanumeric", ISS_at_MMdx ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)
ch_demo
```
<br>
<br>

***

# IV. CH & RaceEthnicity
### PFS
```{r, fig.height = 7}
raceeth_surv <- patients_clean_tumor

mysurv <- Surv(time = raceeth_surv$month_at_progression_drug, event = raceeth_surv$drug_progression_event)
myplot <- survfit(mysurv~raceeth + CH_status, data = raceeth_surv)
# tiff(paste0(path, "/Disparities manuscript/PFS CH raceeth.tiff"))

ggsurvplot(myplot, data = raceeth_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "raceeth",
           linetype = "CH_status",
           palette = c("red", "blue", "#03051AFF"),
           pval = TRUE,
           conf.int = FALSE,

           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 3), linetype = guide_legend(nrow = 2))
# dev.off()
```

```{r}
tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_progression_drug, 
                             event = raceeth_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_progression_drug, 
             event = raceeth_surv$drug_progression_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1 + CH_status, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

### OS
```{r, fig.height = 7}
mysurv <- Surv(time = raceeth_surv$month_at_os, event = raceeth_surv$os_event)
myplot <- survfit(mysurv~raceeth + CH_status, data = raceeth_surv)
# tiff(paste0(path, "/Disparities manuscript/OS CH raceeth.tiff"))

ggsurvplot(myplot, data = raceeth_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "raceeth",
           linetype = "CH_status",
           palette = c("red", "blue", "#03051AFF"),
           pval = TRUE,
           conf.int = FALSE,

           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 3), linetype = guide_legend(nrow = 2))
# dev.off()
```

```{r}
tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_os, 
                             event = raceeth_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_os, 
                   event = raceeth_surv$os_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1 + CH_status, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
<br>
<br>


***

<br>
<br>

# Stratified by raceeth
## WNH
<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
nhis_surv <- patients_clean_tumor %>% 
  filter(raceeth == "White Non-Hispanic")

mysurv <- Surv(time = nhis_surv$month_at_os, event = nhis_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = nhis_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
# tiff(paste0(path, "/Disparities manuscript/OS NHW.tiff"))

ggsurvplot(myplot, data = nhis_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "#03051AFF",
           pval = TRUE,
           conf.int = FALSE,

           # Censor
           censor = TRUE
)
# dev.off()

tbl1 <- nhis_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = nhis_surv$month_at_os, 
                             event = nhis_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = nhis_surv$month_at_os, 
             event = nhis_surv$os_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1, data =  nhis_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_os <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = nhis_surv$month_at_progression_drug, event = nhis_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = nhis_surv)
ggsurvplot(myplot, data = nhis_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "#03051AFF",
           pval = TRUE,
           conf.int = FALSE,
           xlim = c(-15, 250),

           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

tbl1 <- nhis_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = nhis_surv$month_at_progression_drug, 
                             event = nhis_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = nhis_surv$month_at_progression_drug, 
             event = nhis_surv$drug_progression_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1, data =  nhis_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_pfs <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_pfs
```
</div>
</div>

<br>
<br>

## Bl
<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
bl_surv <- patients_clean_tumor %>% 
  filter(raceeth == "Black")

mysurv <- Surv(time = bl_surv$month_at_os, event = bl_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = bl_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
# tiff(paste0(path, "/Disparities manuscript/OS Bl.tiff"))

ggsurvplot(myplot, data = bl_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "red",
           pval = TRUE,
           conf.int = FALSE,
 
           # Censor
           censor = TRUE
)
# dev.off()

tbl1 <- bl_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = bl_surv$month_at_os, 
                             event = bl_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = bl_surv$month_at_os, 
             event = bl_surv$os_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1, data =  bl_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_os <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = bl_surv$month_at_progression_drug, event = bl_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = bl_surv)
ggsurvplot(myplot, data = bl_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "red",
           pval = TRUE,
           conf.int = FALSE,
           xlim = c(-15, 250),

           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

tbl1 <- bl_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = bl_surv$month_at_progression_drug, 
                             event = bl_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = bl_surv$month_at_progression_drug, 
             event = bl_surv$drug_progression_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1, data =  bl_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_pfs <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_pfs
```
</div>
</div>

<br>
<br>

## H

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
his_surv <- patients_clean_tumor %>% 
  filter(raceeth == "Hispanic")

mysurv <- Surv(time = his_surv$month_at_os, event = his_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = his_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
# tiff(paste0(path, "/Disparities manuscript/OS His.tiff"))

ggsurvplot(myplot, data = his_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "blue",
           pval = TRUE,
           conf.int = FALSE,

           # Censor
           censor = TRUE
)
# dev.off()

tbl1 <- his_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = his_surv$month_at_os, 
                             event = his_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = his_surv$month_at_os, 
             event = his_surv$os_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1, data =  his_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_os <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = his_surv$month_at_progression_drug, event = his_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = his_surv)
ggsurvplot(myplot, data = his_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "blue",
           conf.int = FALSE,
           xlim = c(-15, 100),

           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

tbl1 <- his_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = his_surv$month_at_progression_drug, 
                             event = his_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = his_surv$month_at_progression_drug, 
             event = his_surv$drug_progression_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1, data =  his_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_pfs <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_pfs
```
</div>
</div>

<br>
<br>

***

<br>

<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">Sensitivity analysis ISS removed "Unknown" category</span>

</div>
<br>

***
<br>

# I. MM Avatar data description

```{r load2}
germline_patient_data <- readRDS("/Users/colinccm/Documents/GitHub/CHIP-Avatar/germline_patient_data.rds")
path <- ("/Users/colinccm/Documents/GitHub/Gillis/no vpn data/CHIP in Avatar")

germline_patient_data <- germline_patient_data %>% 
  rename(Sex = Gender) %>% 
  filter(is_patient_MM == "Yes" & is_MMDx_close_to_blood == "Yes") %>% 
  distinct(avatar_id, .keep_all = TRUE) %>% 
  filter(raceeth != "Others") %>% 
  mutate(raceeth = factor(raceeth, levels=c("White Non-Hispanic", "Hispanic", "Black"))) %>% 
  mutate(ISS_at_MMdx = replace_na(ISS_at_MMdx, "Unknown")) %>%
  mutate(regimen_categoryVCD_1 = factor(regimen_categoryVCD_1, levels = c(
    "PI + IMIDs", "PI", "IMIDs", "Others"
  )))
```

```{r}
germline_patient_data <- germline_patient_data %>% 
  filter(ISS_at_MMdx != "Unknown") %>% 
  mutate(HCT_comparison = case_when(
    is.na(date_of_bmt_1)              ~ "No HCT",
    time_to_hct < 314                 ~ "Early HCT",
    time_to_hct >= 314                ~ "Late HCT"
  )) %>% 
  mutate(HCT_comparison = factor(HCT_comparison, levels = c("No HCT", "Early HCT", "Late HCT")))

contamination_tumor_ids <- paste("A025767", "A025771", "A007368", "A007322", "A000111", sep = "|")
```

<div class = "row">
<div class = "col-md-5">
```{r}
germ_data <- germline_patient_data %>% 
  mutate(Whole = "MM patients with WES") %>% 
  select(Age_at_MMonly_diagnosis, Sex, raceeth, Whole, ISS_at_MMdx) %>%
  tbl_summary(by = Whole, 
              label = list(raceeth ~ "Race/ethnicity", Age_at_MMonly_diagnosis ~ "Age at MM diagnosis"),
              sort = list(everything() ~ "frequency", ISS_at_MMdx ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2)#,
              # missing = "no"
              ) %>% 
  bold_labels()
germ_data
```
</div>

<div class = "col-md-7">
```{r}
p <- qplot(x =Age_at_MMonly_diagnosis, data=subset(germline_patient_data,!is.na(Age_at_MMonly_diagnosis)), fill=..count.., geom="histogram") 
p + scale_fill_viridis_c(
  alpha = 1,
  begin = 0,
  end = 1,
  direction = 1,
  option = "D",
  values = NULL,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
) +
  theme_minimal(base_size = 16) +
  labs(x="Age at Diagnosis", y="Number of Patient", title="Repartition of Age at Diagnosis in MM Avatar")
```
</div>
</div>

<br>

Patients represented in MM Avatar data are characterized by a median age at diagnosis of `r inline_text(germ_data, variable = Age_at_MMonly_diagnosis, column = "MM patients with WES", pattern = "{median} years old")`. We saw slightly more male `r inline_text(germ_data, variable = Sex, level = "Male", column = "MM patients with WES")`, a majority of White Non-Hispanic `r inline_text(germ_data, variable = raceeth, level = "White Non-Hispanic", column = "MM patients with WES")`(Blacks `r inline_text(germ_data, variable = raceeth, level = "Black", column = "MM patients with WES")`, Hispanic `r inline_text(germ_data, variable = raceeth, level = "Hispanic", column = "MM patients with WES")`).
A higher proportion of patients have a lower clinical stage at blood with `r inline_text(germ_data, variable = ISS_at_MMdx, level = "I", column = "MM patients with WES")` patients with ISS I.

<br>
<br>

***

# II. Age Distribution in Disparities

## Raceeth
```{r}
# tiff(paste0(path, "/Disparities manuscript/age distribution boxplot raceeth.tiff"))
p <- ggplot(germline_patient_data, 
            aes(x=raceeth, y=Age_at_MMonly_diagnosis)) + 
  geom_boxplot(color = c("#03051AFF", "blue", "red"),
               alpha = 0.3,
               fill = c("#03051AFF", "blue", "red")) + 
  theme_minimal(base_size = 23) +
  labs(x=NULL, y="Age at Diagnosis", title="Age of MM Diagnosis Distribution")
p + geom_jitter(shape=16, position=position_jitter(0.2)) +
  stat_compare_means(label.x = 1.75)
# dev.off()
```

```{r}
raceeth_age <- germline_patient_data %>% 
  select(raceeth, Age_at_MMonly_diagnosis) %>%
  tbl_summary(by = raceeth, 
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)
```
<br>

Interestingly, we noticed a significant difference between disparities on the age at diagnosis. White Hispanics are diagnosed at a median age of `r inline_text(raceeth_age, variable = Age_at_MMonly_diagnosis, column = "White Non-Hispanic", pattern = "{median} years old")`. A significant earlier age at diagnosis is measured for Hispanics, `r inline_text(raceeth_age, variable = Age_at_MMonly_diagnosis, column = "Hispanic", pattern = "{median} years old")` , and for Blacks, `r inline_text(raceeth_age, variable = Age_at_MMonly_diagnosis, column = "Black", pattern = "{median} years old")`, compare to Whites Hispanics (`r inline_text(raceeth_age, variable = Age_at_MMonly_diagnosis, column = "p.value")`).

<br>
<br>

***

# III. RaceEthnicity

## Clinical

The p val for the regimen because R is running out of space.
If I run it by hand with `fisher.test(tbl1$raceeth, tbl1$regimen_category_1)`, I have the same.  
I can increase the wrospace `fisher.test(tbl1$raceeth, tbl1$regimen_category_1, workspace=2e9)`which give a good result, we should use that.
But it is a problem in the `gt summary` package which doesn't compute it even if I increse the workspace.
I am going to ask the person who build the package but for now I will give you the p val on the side.

```{r}
# doParallel::registerDoParallel()
tbl1 <-
  germline_patient_data %>% 
  select(Age_at_MMonly_diagnosis, Sex, raceeth, ISS_at_MMdx,
         CH_status,
         HCT_ever, regimen_categoryVCD_1) %>%
  tbl_summary(by = raceeth, 
              label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis",
                           ISS_at_MMdx ~ "ISS at MM diagnosis",
                           HCT_ever ~ "HCT ever",
                           CH_status ~ "CH status"),
              sort = list(everything() ~ "frequency", ISS_at_MMdx ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2)
              ) %>% 
  bold_labels() %>% add_p(test.args = regimen_categoryVCD_1 ~ list(workspace=2e9)) %>% bold_p(t = .05)
```


```{r}
tbl2 <- germline_patient_data %>% 
  select(raceeth,
         time_to_drug, time_to_hct) %>%
  tbl_summary(by = raceeth, 
              label = list(time_to_drug ~ "Time to drug (days)", time_to_hct ~ "Time to HCT (days)"),
              sort = list(everything() ~ "frequency"),
              missing = "no"
              ) %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)

tbl_stack(list(tbl1, tbl2))
```

```{r}
# germline_patient_data %>% 
#   select(time_to_drug, HCT_comparison, raceeth) %>%
#   tbl_summary(by = raceeth, 
#               label = list(time_to_drug ~ "Time to drug (days)", HCT_comparison ~ "Time to HCT (days)"),
#               sort = list(everything() ~ "frequency"),
#               missing = "no") %>% bold_labels() %>% add_overall() %>% add_p() %>% bold_p(t = .05)
```

<br>
<br>

## Survivals

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
raceeth_surv <- germline_patient_data

mysurv <- Surv(time = raceeth_surv$month_at_os, event = raceeth_surv$os_event)
myplot <- survfit(mysurv~raceeth, data = raceeth_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
# pdf(paste0(path, "/Disparities manuscript/OS raceeth.pdf"))
ggsurvplot(myplot, data = raceeth_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           legend.labs = c("White Non-Hispanic", "Hispanic", "Black"),
           palette = c("#03051AFF", "blue", "red"),
           pval = TRUE,
           conf.int = FALSE,
           # xlim = c(-15, 400),


           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3))
# dev.off()

tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_os, 
                             event = raceeth_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_os, 
             event = raceeth_surv$os_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# eth_cox_os <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
# eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = raceeth_surv$month_at_progression_drug, event = raceeth_surv$drug_progression_event)
myplot <- survfit(mysurv~raceeth, data = raceeth_surv)
# pdf(paste0(path, "/Disparities manuscript/PFS raceeth.pdf"))

ggsurvplot(myplot, data = raceeth_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           legend.labs = c("White Non-Hispanic", "Hispanic", "Black"),
           palette = c("#03051AFF", "blue", "red"),
           pval = TRUE,
           conf.int = FALSE,
           # xlim = c(-10, 250),


           # Censor
           censor = TRUE
) + guides(colour = guide_legend(nrow = 3))
# dev.off()

tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_progression_drug, 
                             event = raceeth_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_progression_drug, 
             event = raceeth_surv$drug_progression_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# eth_cox_pfs <- 
  tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
# eth_cox_pfs
```
</div>
</div>

<br>
<br>

***

# IV. CH

From here, remove 5 patients with tumor contamination

```{r}
patients_clean_tumor <- germline_patient_data %>% 
  filter(!str_detect(avatar_id, contamination_tumor_ids))
```

```{r}
ch_char <- patients_clean_tumor %>% 
  mutate(Whole = "MM patients at time of blood with WES") %>% 
  select(Whole, CH_status) %>%
  tbl_summary(by = Whole, 
              label = list(CH_status ~ "CH status"),
              sort = list(everything() ~ "frequency"),
              missing = "no") %>% 
  bold_labels()
ch_char
```
<br>
<br>

Whole Exome Sequencing data identified CH mutations in `r inline_text(ch_char, variable = CH_status, level = "CH", column = "MM patients at time of blood with WES")`.
+ Add bioinfo results (oncoplot?).  
Type of genes affected...

<br>
<br>

```{r}
# ch_samples <- patients_clean_tumor %>%
#   select(CH_status) %>% 
#   tbl_summary(sort = list(everything() ~ "frequency")) %>% 
#   bold_labels()
# ch_samples
patients_clean_tumor %>%
  select(Age_at_MMonly_diagnosis, 
         Sex, raceeth, ISS_at_MMdx, ISS_at_MMdx,
         CH_status) %>% 
  tbl_summary(by = CH_status, 
              label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis", CH_status ~ "CH status"),
              sort = list(everything() ~ "frequency", ISS_at_MMdx ~ "alphanumeric", ISS_at_MMdx ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              percent = "row",
              missing = "no") %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)

ch_demo <- patients_clean_tumor %>%
  select(Age_at_MMonly_diagnosis, 
         Sex, raceeth, ISS_at_MMdx, ISS_at_MMdx,
         CH_status) %>% 
  tbl_summary(by = CH_status, 
              label = list(Age_at_MMonly_diagnosis ~ "Age at MM diagnosis", CH_status ~ "CH status"),
              sort = list(everything() ~ "frequency", ISS_at_MMdx ~ "alphanumeric", ISS_at_MMdx ~ "alphanumeric"),
              digits = list(c(Age_at_MMonly_diagnosis) ~ 2),
              missing = "no") %>% 
  bold_labels() %>% add_p() %>% bold_p(t = .05)
ch_demo
```
<br>
<br>

***

# IV. CH & RaceEthnicity
### PFS
```{r, fig.height = 7}
raceeth_surv <- patients_clean_tumor

mysurv <- Surv(time = raceeth_surv$month_at_progression_drug, event = raceeth_surv$drug_progression_event)
myplot <- survfit(mysurv~raceeth + CH_status, data = raceeth_surv)
# tiff(paste0(path, "/Disparities manuscript/PFS CH raceeth.tiff"))

ggsurvplot(myplot, data = raceeth_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "raceeth",
           linetype = "CH_status",
           palette = c("red", "blue", "#03051AFF"),
           pval = TRUE,
           conf.int = FALSE,

           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 3), linetype = guide_legend(nrow = 2))
# dev.off()
```

```{r}
tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_progression_drug, 
                             event = raceeth_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_progression_drug, 
             event = raceeth_surv$drug_progression_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1 + CH_status, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

### OS
```{r, fig.height = 7}
mysurv <- Surv(time = raceeth_surv$month_at_os, event = raceeth_surv$os_event)
myplot <- survfit(mysurv~raceeth + CH_status, data = raceeth_surv)
# tiff(paste0(path, "/Disparities manuscript/OS CH raceeth.tiff"))

ggsurvplot(myplot, data = raceeth_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           color = "raceeth",
           linetype = "CH_status",
           palette = c("red", "blue", "#03051AFF"),
           pval = TRUE,
           conf.int = FALSE,

           # Censor
           censor = TRUE
)+ guides(colour = guide_legend(nrow = 3), linetype = guide_legend(nrow = 2))
# dev.off()
```

```{r}
tbl1 <- raceeth_surv %>% select(raceeth, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1, CH_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = raceeth_surv$month_at_os, 
                             event = raceeth_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = raceeth_surv$month_at_os, 
                   event = raceeth_surv$os_event) ~ raceeth + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1 + CH_status, data =  raceeth_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
<br>
<br>


***

<br>
<br>

# Stratified by raceeth
## WNH
<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
nhis_surv <- patients_clean_tumor %>% 
  filter(raceeth == "White Non-Hispanic")

mysurv <- Surv(time = nhis_surv$month_at_os, event = nhis_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = nhis_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
# tiff(paste0(path, "/Disparities manuscript/OS NHW.tiff"))

ggsurvplot(myplot, data = nhis_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "#03051AFF",
           pval = TRUE,
           conf.int = FALSE,

           # Censor
           censor = TRUE
)
# dev.off()

tbl1 <- nhis_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = nhis_surv$month_at_os, 
                             event = nhis_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = nhis_surv$month_at_os, 
             event = nhis_surv$os_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1, data =  nhis_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_os <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = nhis_surv$month_at_progression_drug, event = nhis_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = nhis_surv)
ggsurvplot(myplot, data = nhis_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "#03051AFF",
           pval = TRUE,
           conf.int = FALSE,
           xlim = c(-15, 250),

           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

tbl1 <- nhis_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = nhis_surv$month_at_progression_drug, 
                             event = nhis_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = nhis_surv$month_at_progression_drug, 
             event = nhis_surv$drug_progression_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1, data =  nhis_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_pfs <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_pfs
```
</div>
</div>

<br>
<br>

## Bl
<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
bl_surv <- patients_clean_tumor %>% 
  filter(raceeth == "Black")

mysurv <- Surv(time = bl_surv$month_at_os, event = bl_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = bl_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
# tiff(paste0(path, "/Disparities manuscript/OS Bl.tiff"))

ggsurvplot(myplot, data = bl_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "red",
           pval = TRUE,
           conf.int = FALSE,
 
           # Censor
           censor = TRUE
)
# dev.off()

tbl1 <- bl_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = bl_surv$month_at_os, 
                             event = bl_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = bl_surv$month_at_os, 
             event = bl_surv$os_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1, data =  bl_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_os <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = bl_surv$month_at_progression_drug, event = bl_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = bl_surv)
ggsurvplot(myplot, data = bl_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "red",
           pval = TRUE,
           conf.int = FALSE,
           xlim = c(-15, 250),

           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

tbl1 <- bl_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = bl_surv$month_at_progression_drug, 
                             event = bl_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = bl_surv$month_at_progression_drug, 
             event = bl_surv$drug_progression_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1, data =  bl_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_pfs <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_pfs
```
</div>
</div>

<br>
<br>

## H

<div class = "row">
<div class = "col-md-6">
```{r, fig.height = 7}
his_surv <- patients_clean_tumor %>% 
  filter(raceeth == "Hispanic")

mysurv <- Surv(time = his_surv$month_at_os, event = his_surv$os_event)
myplot <- survfit(mysurv~CH_status, data = his_surv)
# jpeg(paste0(path, "/Figures/Survivals/Demographic/OS Ethnicity.jpeg"), width = 1200, height = 900)
# tiff(paste0(path, "/Disparities manuscript/OS His.tiff"))

ggsurvplot(myplot, data = his_surv,
           title = "OS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "blue",
           pval = TRUE,
           conf.int = FALSE,

           # Censor
           censor = TRUE
)
# dev.off()

tbl1 <- his_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = his_surv$month_at_os, 
                             event = his_surv$os_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = his_surv$month_at_os, 
             event = his_surv$os_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + HCT_ever + regimen_categoryVCD_1, data =  his_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_os <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_os
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
mysurv <- Surv(time = his_surv$month_at_progression_drug, event = his_surv$drug_progression_event)
myplot <- survfit(mysurv~CH_status, data = his_surv)
ggsurvplot(myplot, data = his_surv,
           title = "PFS",
           font.main = c(24, "bold", "black"),
           font.x = c(20, "bold", "black"),
           font.y = c(20, "bold", "black"),
           font.legend = c(20, "bold", "black"),
           font.tickslab = c(18, "bold", "black"),
           size = 1.5,
           
           xlab = "Time in months", 
           legend = "top",
           legend.title = "",
           # # legend.labs = c("Hispanic", "Non-Hispanic"),
           linetype = "CH_status",
           color = "blue",
           conf.int = FALSE,
           xlim = c(-15, 100),

           # Add risk table
           tables.height = 0.3,
           risk.table.title = "Risk table (number(%))",
           risk.table = "abs_pct",
           risk.table.y.text = FALSE,
           risk.table.fontsize = 4,
           tables.theme = theme_survminer(base_size = 5,
                                          font.main = c(16, "bold", "black"),
                                          font.x = c(16, "bold", "black"),
                                          font.y = c(16, "bold", "transparent"),
                                          font.tickslab = c(19, "bold", "black")
           ),
           # Censor
           censor = TRUE
)

tbl1 <- his_surv %>% select(CH_status, Age_at_MMonly_diagnosis, Sex, ISS_at_MMdx, HCT_ever, regimen_categoryVCD_1) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = his_surv$month_at_progression_drug, 
                             event = his_surv$drug_progression_event)),
                   exponentiate = TRUE) %>% bold_p(t = .05) %>% add_nevent() %>% 
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = his_surv$month_at_progression_drug, 
             event = his_surv$drug_progression_event) ~ CH_status + Age_at_MMonly_diagnosis + Sex + ISS_at_MMdx + strata(HCT_ever) + regimen_categoryVCD_1, data =  his_surv) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
eth_cox_pfs <- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
eth_cox_pfs
```
</div>
</div>

<br>






